---
title: "Additional microbiome analyses for _Environmental bacteria link green schoolyards to better child well-being and fewer allergy symptoms_"
author: "Wenke Smets, Hans Van Calster"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(knitr)
library(here)
opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6
)
opts_knit$set(root.dir = here())
library(frictionless)
library(tidytacos)
library(radEmu)
library(glmmTMB)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(forcats)
library(lme4)

conflicted::conflicts_prefer(dplyr::select)
conflicted::conflicts_prefer(dplyr::filter)

"%!in%" <- function(x, y) !("%in%"(x, y))

# Copied from tidytacos:::tidy_count_to_matrix (internal, not exported)
# tidytacos version: 1.0.9
tidy_count_to_matrix <- function(counts, value = count) {
  if (!is.data.frame(counts) || is.null(counts$taxon_id) || is.null(counts$sample_id)) {
    stop("first argument should be a counts table (data frame)")
  }
  value <- rlang::enquo(value)
  counts_wide <- counts %>%
    dplyr::select(sample_id, taxon_id, !!value) %>%
    tidyr::spread(key = taxon_id, value = !!value, fill = 0)
  counts_wide %>%
    dplyr::select(-sample_id) %>%
    as.matrix() %>%
    `row.names<-`(counts_wide$sample_id)
}
```

```{r cache=TRUE, results='hide'}
1 + 1
# just a trick to prevent the folder containing figures from being removed
# when knitting a self-contained document
```

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}

.warning {
  background-color: #fcf8e3;
  color: #8a6d3b;
  padding: 15px;
  border-left: 5px solid #faebcc;
  border-radius: 4px;
  margin-bottom: 15px;
}
```

::: warning
**Warning:** This document contains the analysis pipeline that was used for the paper *Environmental bacteria link green schoolyards to better child well-being and fewer allergy symptoms*.
While the code is identical, the data used in this document (513 participants) is a (large) subset of the data used in the paper (529 participants).
The reason for this is that some participants / parents did not give consent for long-term storage / archiving of their data.
As such, we had permission to use their data in publications, but not to publish their data in a long-term public archive.
The data used in this repository, which are downloaded from <https://doi.org/10.5281/zenodo.10527033> (see `source/reproduce.R`) [@vancalster2024], therefore only include data from participants / parents that gave consent for long-term storage of their data.
As a consequence, the results from the paper are not exactly reproducible.
:::

# Read resources

```{r read-package}
cleaned_data <- read_package(
  here::here(
    "data",
    "datapackage_an_zenodo_cleaned_data",
    "datapackage.json"
  )
)

derived_data <- read_package(
  here::here(
    "data",
    "datapackage_an_zenodo_derived_data",
    "datapackage.json"
  )
)
```

```{r which-resources}
resource_names(cleaned_data)
resource_names(derived_data)
```


# Join landscape naturalness and greenness information to microbiome sample data

```{r}
landscape_data <- read_resource(
  package = cleaned_data,
  resource_name = "wp1_landscape_level_data"
)

# glimpse(landscape_data)

school_data <- read_resource(
  package = derived_data,
  resource_name = "wp1_school_level_key_variables"
)

plgr_biodiv <- read_resource(
  package = cleaned_data,
  resource_name = "wp3_ucl_playground_biodiversity"
)

greenness <- read_resource(
  cleaned_data,
  "wp3_ucl_greenness_indicators"
)

plgr_areas <- greenness %>%
  filter(
    indicator == "greenness2",
    buffer_size == 0,
    area_considered == "playground limit"
  ) %>%
  select(code_area, playground_area = buffer_area)

plgr_biodiv_2 <- plgr_biodiv %>%
  group_by(school_id) %>%
  arrange(school_id) %>%
  mutate(
    frequency_of_use = if_else(
      n() == 1 & is.na(frequency_of_use),
      "Daily",
      frequency_of_use
    ),
    weight_freq = recode_values(
      frequency_of_use,
      "Daily" ~ 1,
      "Weekly" ~ 1 / 5,
      "Monthly" ~ 1 / 20
    ),
    weight_freq = weight_freq / sum(weight_freq, na.rm = TRUE)
  ) %>%
  inner_join(plgr_areas, by = join_by(code_playground == code_area)) %>%
  mutate(
    weight_area = playground_area / sum(playground_area, na.rm = TRUE),
    weight = weight_area * weight_freq /
      sum(weight_area * weight_freq, na.rm = TRUE)
  )

school_biodiv_derived <- plgr_biodiv_2 %>%
  ungroup() %>%
  mutate(
    tree_number_rescaled =
      (a_tree_number - min(a_tree_number, na.rm = TRUE)) /
        (max(a_tree_number, na.rm = TRUE) - min(a_tree_number, na.rm = TRUE))
  ) %>%
  group_by(school_id) %>%
  summarise(
    water_feature = sum(weight * water_feature),
    vegetation = sum(weight * vegetation),
    sand = sum(weight * sand),
    woodchips = sum(weight * woodchips),
    gravels = sum(weight * gravels),
    b_edge_shrub = sum(weight * b_edge_shrub),
    c_grass = sum(weight * c_grass),
    d_flowerbed = sum(weight * d_flowerbed),
    e_weeds = sum(weight * e_weeds),
    f_flowerbox_gardenbox = sum(weight * f_flowerbox_gardenbox),
    g_garden = sum(weight * g_garden),
    h_treestump = sum(weight * h_treestump),
    i_greenwall = sum(weight * i_greenwall),
    j_branches = sum(weight * j_branches),
    k_greenroof = sum(weight * k_greenroof),
    tree_number_rescaled = sum(weight * tree_number_rescaled^(1 / 5))
  )


# glimpse(envi_mb_data)

envdat <- landscape_data %>%
  select(school_id, landscape_type, no2, bc, pm10 = pm10_min_pm25) %>%
  left_join(school_biodiv_derived) %>%
  left_join(
    school_data %>%
      mutate(playgreen = playground__area_green / playground__area) %>%
      mutate(extragreen = extra_school__area_green / extra_school__area) %>%
      mutate(outgreen100 = extra_100m__area_green / extra_100m__area) %>%
      mutate(outgreen300 = extra_300m__area_green / extra_300m__area) %>%
      select(
        school_id, community, school_context, allergenic_potential, playgreen,
        extragreen, outgreen100, outgreen300, sirm = mean_sirm_logratio, pm25,
        school_ses = ses, natural_elements_score, tree_number, province_name,
        case_control_id
      )
  )

rm(landscape_data, school_data, plgr_areas, plgr_biodiv,
   plgr_biodiv_2, school_biodiv_derived, greenness)
```


# Create environmental `tidtytacos` object

Let's make a new `tidytacos` object from the samples, taxa and counts tables for the environmental microbiomes that includes the naturalness of the landscape.

```{r}
envir_microbiome <- read_tidytacos(here("data", "micdata", "environ"))

envir_microbiome$samples <- envir_microbiome$samples |>
  left_join(envdat)

str(envir_microbiome, 1)
```


# Preprocess environmental microbiome data

```{r}
unique(envir_microbiome$samples$sample_type)
envir_microbiome$samples <- envir_microbiome$samples %>%
  mutate(
    sample_type = ifelse(
      sample_type %in% c("sanddirtdust", "dirt/dust"),
      "dirt",
      sample_type)
  )
```


# Join participant information to skin microbiome sample data

First joining several child characteristics with microbiome data.
Some doubles occur in the questionnaires: mean of numeric responses per child.

```{r}
child_data1 <- read_resource(
  package = derived_data,
  resource_name = "wp1_child_level_key_variables"
)
child_data2 <- read_resource(
  package = cleaned_data,
  resource_name = "wp5_parents_outdoor_play_data_part2"
)

child_data2 <- child_data2 %>%
  filter(
    question_code %in%
      c("x5_5", "x5_7", "x5_8", "x5_9", "x5_11", "x5_12", "x5_13", "x5_14",
        "x5_15", "x5_17", "x5_18")
  ) %>%
  mutate(response = ifelse(
    response == "no association - association",
    0.5,
    ifelse(
      response == "association",
      1,
      ifelse(response == "no association", 0, NA)
      )
    )
  ) %>%
  group_by(child_id, question_code) %>%
  summarise(response = mean(response)) %>%
  ungroup() %>%
  group_by(child_id) %>%
  summarise(nature_exposure = sum(response) / 11)

child_data3 <- read_resource(
  package = cleaned_data,
  resource_name = "wp5_parents_outdoor_play_data_part3"
)

child_data3 <- child_data3 %>%
  filter(question_code %in% c("x5_20", "x5_21", "x5_22", "x5_23")) %>%
  mutate(response = ifelse(response == "< 1", 0.5, ifelse(
    response == "1-2", 1.5, ifelse(response == "2-3", 2.5, ifelse(
      response == "3-4", 3.5, ifelse(
        response == "< 1 - 1-2",
        1,
        ifelse(response == "1-2 - 3-4", 2.5, ifelse(
          response == "0", 0, ifelse(response == "> 4", 5, ifelse(
            response == "1-2 - 2-3", 2, ifelse(response == "3-4 - > 4", 4, ifelse(
              response == "0 - < 1", 0.5, ifelse(response == "2-3 - 3-4", 3, NA)
            ))
          ))
        ))
      )
    ))
  ))) %>%
  group_by(child_id, question_code) %>%
  summarize(response = mean(response)) %>%
  pivot_wider(names_from = question_code, values_from = response) %>%
  mutate(sitsindoors = ((x5_20 + x5_22) * 5 + (x5_21 + x5_23) * 2) / 7) %>%
  select(child_id, sitsindoors)


child_data4 <- read_resource(
  package = cleaned_data,
  resource_name = "wp5_parents_living_environment_data"
)

garden_data_recoded <- child_data4 %>%
  filter(
    question_short %in%
      c("garden_present", "garden_type", "garden_usage", "garden_size")
  ) %>%
  mutate(response_recoded = case_when(
    question_short == "garden_present" & response_values == "1" ~ 0,
    question_short == "garden_present" & response_values != "1" ~ 1,
    question_short == "garden_size" & response_values == "1" ~ 0,
    question_short == "garden_size" & response_values == "2" ~ 1,
    question_short == "garden_size" & response_values == "2 - 3" ~ 1.5,
    question_short == "garden_size" & response_values == "3" ~ 2,
    question_short == "garden_size" & response_values == "4" ~ 3,
    question_short == "garden_type" & response_values == "1" ~ 0,
    question_short == "garden_type" & response_values == "1 - 3" ~ 1,
    question_short == "garden_type" & response_values == "2" ~ 1,
    question_short == "garden_type" & response_values == "3" ~ 2,
    question_short == "garden_type" & response_values == "4" ~ 1.5,
    question_short == "garden_usage" ~ as.numeric(response_values) - 1
  ))

garden_data <- garden_data_recoded %>%
  pivot_wider(
    id_cols = c(child_id, replicate_id, case),
    names_from = question_short,
    values_from = response_recoded,
    values_fill = list(response_recoded = 0)
  ) %>%
  mutate(
    garden_score = (garden_size / 3 + garden_type / 2 + garden_usage / 3) / 3
  ) %>%
  select(child_id, garden_present, garden_score)

kiddat <- child_data1 %>%
  left_join(child_data2) %>%
  left_join(child_data3) %>%
  left_join(garden_data)

kiddat_num <- unique(kiddat) %>%
  group_by(child_id) %>%
  summarise(across(where(is.numeric), mean))
kiddat_chr <- kiddat %>%
  select(child_id, school_id, class_id, biological_sex, family_situation,
         origin_simplified) %>%
  unique()
kiddat <- left_join(kiddat_num, kiddat_chr)

rm(child_data1, child_data2, child_data3, garden_data, garden_data_recoded,
   child_data4, kiddat_num, kiddat_chr)
```


# Create skin `tidytacos` object

```{r skin-microbiome}
skin_microbiome <- read_tidytacos(here("data", "micdata", "skin"))

skin_microbiome$samples <- skin_microbiome$samples %>%
  left_join(kiddat, by = join_by(school_id, class_id, child_id))

str(skin_microbiome, 1)
```

# Data prep

```{r}
# Data per child:
cdata <- kiddat %>%
  select(
    school_id, child_id, biological_sex, nature_exposure, nature_connectedness,
    sitsindoors, ses, family_situation, origin_simplified, t_score_sch,
    t_score_par, t_score_soc, t_score_phy, t_score_pwb, eczema, rhinitis,
    wheezing
  )

# Data per school:
sdata <- envdat %>%
  select(
    school_id, school_context, allergenic_potential, playgreen, extragreen,
    outgreen100, outgreen300, landscape_type, no2, bc, pm10, sirm, pm25,
    school_ses, community, tree_number, natural_elements_score
  )

# Merge child and school data
kiddatnew <- left_join(cdata, sdata, by = join_by(school_id))

# Skin microbiome plus metadata
runs <- skin_microbiome
runs$samples <- runs$samples %>%
  left_join(sdata)
```


# Cleaning and setting up microbiome data

```{r}
# remove low-read samples§
rune <- envir_microbiome %>%
  filter_samples(lib_size > 350) %>%
  mutate_samples(sample = paste(school_id, sample_type, lib_size)) %>%
  mutate_samples(scale_mean_log_sirm = as.numeric(scale(sirm)))
runs <- runs %>%
  filter_samples(lib_size > 600) %>%
  mutate_samples(sample = child_id) %>%
  mutate_samples(scale_mean_log_sirm = as.numeric(scale(sirm))) %>%
  add_total_count()

# adding div of most abundant skin classes

scl <- runs %>%
  aggregate_taxa(rank = "class") %>%
  add_mean_rel_abundance() %>%
  taxa() %>%
  arrange(desc(mean_rel_abundance)) %>%
  select(class)
classes <- scl$class[1:11][!is.na(scl$class[1:11])] # removing class NA, keeping ten most abundant classes

set.seed(2487)
r <- runs %>%
  add_mean_rel_abundance() %>%
  select_samples(-inverse_simpson) %>%
  add_alphas(method = c("obs", "invsimpson"))
sdiv <- r$samples %>%
  select(child_id, school_id, skin_richness = obs, skin_simpson = invsimpson)
runs$samples <- left_join(runs$samples, sdiv)

namesrich <- c()
namessimp <- c()

for (c in classes) {
  r <- runs %>%
    filter_taxa(class == c) %>%
    add_alphas(method = c("obs", "invsimpson"))
  sdivs <- r$samples %>%
    select(child_id, obs, invsimpson) %>%
    dplyr::rename(!!sym(paste(str_sub(c, 1, 5), "richness", sep = "_")) := obs, !!sym(paste(str_sub(c, 1, 5), "simpson", sep = "_")) := invsimpson)
  runs$samples <- left_join(runs$samples, sdivs)
  namesrich <- append(namesrich, paste(str_sub(c, 1, 5), "richness", sep = "_"))
  namessimp <- append(namessimp, paste(str_sub(c, 1, 5), "simpson", sep = "_"))
}
classdiv <- tibble(classes, namesrich, namessimp)
runs$samples <- runs$samples %>%
  left_join(unique(select(rune$samples, school_id)))

tacosum(rune)
runeg <- aggregate_taxa(rune, rank = "genus") %>%
  add_mean_rel_abundance()
tacosum(runeg)
tacosum(runs)
runsg <- aggregate_taxa(runs, rank = "genus") %>%
  add_mean_rel_abundance()
tacosum(runsg)

rm(r, sdivs, sdiv, scl)
```


# Allergy data reformatting

```{r}
runs <- runs %>%
  mutate_samples(sample = child_id) %>%
  mutate_samples(Eczema = ifelse(eczema < 0, 0, 1)) %>%
  mutate_samples(Wheezing = ifelse(wheezing < 0, 0, 1)) %>%
  mutate_samples(Rhinitis = ifelse(rhinitis < 0, 0, 1)) %>%
  mutate_samples(allergy_symptoms = Eczema + Wheezing + Rhinitis)


# allergies per school, with omission of schools with only 1 or 2 pupils

allergies <- runs$samples %>%
  drop_na(allergy_symptoms) %>%
  group_by(school_id) %>%
  summarise(
    meaneczema = mean(Eczema),
    meanwheezing = mean(Wheezing),
    meanrhinitis = mean(Rhinitis),
    allergy_symptoms = mean(allergy_symptoms),
    ne = sum(!is.na(Eczema))) %>%
  mutate(meaneczema = ifelse(ne < 3, NA, meaneczema)) %>%
  drop_na(meaneczema) %>%
  select(-ne)

rune$samples <- left_join(rune$samples, allergies)
data <- runs$samples


# means of allergic symptoms

dat <- data %>%
  select(Eczema, nature_exposure) %>%
  drop_na(Eczema)
sum(dat$Eczema) / nrow(dat)
dat <- data %>%
  select(Wheezing, nature_exposure) %>%
  drop_na(Wheezing)
sum(dat$Wheezing) / nrow(dat)
dat <- data %>%
  select(Rhinitis, nature_exposure) %>%
  drop_na(Rhinitis)
sum(dat$Rhinitis) / nrow(dat)
```


# Allergy-specific bacteria in plants (radEmu)

Unify ASV names across environmental and skin taxa:

```{r}
taxn <- runs %>%
  mutate_samples(sample_type = "skin") %>%
  merge_tidytacos(rune, taxon_identifier = taxon) %>%
  select_taxa(-species) %>%
  add_taxon_name() %>%
  add_prevalence() %>%
  filter_taxa(prevalence > 2)
taxn <- taxn$taxa %>%
  select(taxon_name, taxon)
```

Are 100 most prevalent plant taxa related to any of the allergies?
(this code could take a while to run!)

```{r rademu-plant-samples}
# pooling 3 plant samples per school
y1 <- rune %>%
  filter_samples(sample_type == "plant") %>%
  select_samples(
    school_id, sample_id, meaneczema, meanrhinitis, meanwheezing,
    allergy_symptoms, natural_elements_score, school_context, landscape_type) %>%
  aggregate_samples() %>%
  add_prevalence()

ggplot(y1$taxa, aes(x = prevalence)) +
  geom_histogram(binwidth = 1) +
  xlim(c(2, 30)) +
  # ylim(c(1,500))+
  # scale_y_log10()+
  labs(x = "Prevalence in plant samples", y = "ASV count") +
  geom_vline(xintercept = 7.5, col = "blue")

y1$samples <- drop_na(y1$samples, allergy_symptoms)
y1$counts <- filter(y1$counts, sample_id %in% y1$samples$sample_id)

# setting up to select only the 100 most prevalent ASVs
y1$taxa <- y1$taxa %>%
  arrange(desc(prevalence)) %>%
  left_join(taxn) %>%
  mutate(keep = F) %>%
  {
    .$keep[1:100] <- T
    .
  }

y1 <- y1 %>%
  mutate_taxa(name = ifelse(keep == FALSE, "other", taxon_name)) %>%
  select_taxa(taxon_id, name) %>%
  aggregate_taxa() %>%
  add_prevalence(relative = F)

mostprevplant <- y1$taxa %>%
  left_join(rune$taxa) %>%
  select(name, family, class, phylum, prevalence)
#write.csv(mostprevplant, "mostprevplant.csv", quote = F, row.names = F)

ggplot(y1$taxa, aes(x = prevalence)) +
  geom_histogram()

samp <- select(
  y1$samples,
  sample_id,
  meaneczema,
  meanrhinitis,
  meanwheezing,
  allergy_symptoms,
  school_id
)
rownames(samp) <- samp$sample_id
y <- tidy_count_to_matrix(y1$counts)[samp$sample_id, ]

# differential abundance analysis with radEmu
# increase tolerance for more reliable pvalues, at the cost of longer runtime
#radEmu::make_design_matrix(formula = ~ meaneczema, data = samp, Y = y)
xfun::cache_exec(
  {
    fit1 <- radEmu::emuFit(
      formula = ~ meaneczema, data = samp, Y = y,
      test_kj = data.frame(k = "meaneczema", j = seq_len(ncol(y))),
      tolerance = 1e-2
    )
    fit2 <- radEmu::emuFit(
      formula = ~ meanrhinitis, data = samp, Y = y,
      test_kj = data.frame(k = "meanrhinitis", j = seq_len(ncol(y))),
      tolerance = 1e-2
    )
    fit3 <- radEmu::emuFit(
      formula = ~ meanwheezing, data = samp, Y = y,
      test_kj = data.frame(k = "meanwheezing", j = seq_len(ncol(y))),
      tolerance = 1e-2
    )
    fit4 <- radEmu::emuFit(
      formula = ~ allergy_symptoms, data = samp, Y = y,
      test_kj = data.frame(k = "allergy_symptoms", j = seq_len(ncol(y))),
      tolerance = 1e-2
    )
  },
  path = "source/extra_microbiome_analyses_cache/"
)

ecz <- fit1$coef %>%
  filter(covariate == "meaneczema") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
ecz

rhi <- fit2$coef %>%
  filter(covariate == "meanrhinitis") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
rhi

whe <- fit3$coef %>%
  filter(covariate == "meanwheezing") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
whe

sym <- fit4$coef %>%
  filter(covariate == "allergy_symptoms") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
sym

planttaxa <- rbind(ecz, rhi, whe, sym)

sigtax <- planttaxa %>%
  filter(
    pval < 0.05 & covariate %in%
      c("meaneczema", "meanrhinitis", "meanwheezing", "allergy_symptoms")
  )
planttaxa <- planttaxa %>%
  filter(name %in% sigtax$name) %>%
  mutate(significance = ifelse(pval < 0.05, "*", "")) %>%
  mutate(
    covariate = fct_recode(
      covariate,
      "Eczema symptoms" = "meaneczema",
      "Rhinitis symptoms" = "meanrhinitis",
      "Wheezing & asthma symptoms" = "meanwheezing",
      "Allergy categories count" = "allergy_symptoms",
      "Urban landscape" = "landscape_typelow naturalness",
      "Grey schoolground" = "school_contextlow greenness",
      "Natural elements score" = "natural_elements_score"
    )
  ) %>%
  mutate(
    covariate = fct_relevel(
      covariate,
      "Urban landscape",
      "Grey schoolground",
      "Natural elements score",
      "Eczema symptoms",
      "Rhinitis symptoms",
      "Wheezing & asthma symptoms",
      "Allergy categories count"
    )
  )

# Figures

ggplot(planttaxa, aes(y = covariate, x = name, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 5) +
  xlab("Plant ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))


fplanttaxa <- planttaxa %>%
  filter(
    covariate %in% c(
      "Eczema symptoms",
      "Rhinitis symptoms",
      "Wheezing & asthma symptoms",
      "Allergy categories count"
    )
  )
ggplot(fplanttaxa, aes(y = covariate, x = name, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 5) +
  xlab("Plant ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))
```


# Allergy-specific bacteria in sand dirt dust (radEmu)

(Figure in supplementary materials)

```{r rademu-sanddirtdust-samples}
y1 <- rune %>%
  filter_samples(sample_type %in% c("sand", "dirt", "dust")) %>%
  select_samples(
    school_id, sample_id, sample_type, meaneczema, meanrhinitis,
    meanwheezing, allergy_symptoms, natural_elements_score, school_context,
    landscape_type
  ) %>%
  aggregate_samples() %>%
  add_prevalence()

y1$samples <- drop_na(y1$samples, allergy_symptoms)
y1$counts <- filter(y1$counts, sample_id %in% y1$samples$sample_id)

y1$taxa <- y1$taxa %>%
  arrange(desc(prevalence)) %>%
  left_join(taxn) %>%
  mutate(keep = FALSE) %>%
  {
    .$keep[1:100] <- TRUE
    .
  }

y1 <- y1 %>%
  mutate_taxa(name = ifelse(keep == FALSE, "other", taxon_name)) %>%
  select_taxa(taxon_id, name) %>%
  aggregate_taxa()
samp <- select(
  y1$samples, sample_id, sample_type, meaneczema, meanrhinitis, meanwheezing,
  allergy_symptoms, school_id
)
rownames(samp) <- samp$sample_id
y <- tidy_count_to_matrix(y1$counts)[samp$sample_id, ]
design <- radEmu::make_design_matrix(formula = ~ sample_type + meaneczema, data = samp)
colnames(design)

xfun::cache_exec(
  {
    fit1 <- radEmu::emuFit(
      formula = ~ sample_type + meaneczema,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(j = c(1:101), k = 4)
    )
    fit2 <- radEmu::emuFit(
      formula = ~ sample_type + meanrhinitis,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(j = c(1:101), k = 4)
    )
    fit3 <- radEmu::emuFit(
      formula = ~ sample_type + meanwheezing,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(j = c(1:101), k = 4)
    )
    fit4 <- radEmu::emuFit(
      formula = ~ sample_type + allergy_symptoms,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(j = c(1:101), k = 4)
    )
  },
  path = "source/extra_microbiome_analyses_cache/"
)
ecz <- fit1$coef %>%
  filter(covariate == "meaneczema") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
ecz

rhi <- fit2$coef %>%
  filter(covariate == "meanrhinitis") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
rhi

whe <- fit3$coef %>%
  filter(covariate == "meanwheezing") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
whe

sym <- fit4$coef %>%
  filter(covariate == "allergy_symptoms") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
sym

soiltaxa <- rbind(ecz, rhi, whe, sym)

sigtax <- soiltaxa %>%
  filter(
    pval < 0.05 & covariate %in%
      c("meaneczema", "meanrhinitis", "meanwheezing", "allergy_symptoms")
  )
soiltaxa <- soiltaxa %>%
  filter(name %in% sigtax$name) %>%
  mutate(significance = ifelse(pval < 0.05, "*", "")) %>%
  mutate(
    covariate = fct_recode(
      covariate,
      "Eczema symptoms" = "meaneczema",
      "Rhinitis symptoms" = "meanrhinitis",
      "Wheezing & asthma symptoms" = "meanwheezing",
      "Allergy categories count" = "allergy_symptoms",
      "Urban landscape" = "landscape_typelow naturalness",
      "Grey schoolground" = "school_contextlow greenness",
      "Natural elements score" = "natural_elements_score"
    )
  ) %>%
  mutate(
    covariate = fct_relevel(
      covariate,
      "Urban landscape",
      "Grey schoolground",
      "Natural elements score",
      "Eczema symptoms",
      "Rhinitis symptoms",
      "Wheezing & asthma symptoms",
      "Allergy categories count"
    )
  )

ggplot(soiltaxa, aes(y = covariate, x = name, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 5) +
  xlab("Sand/soil ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

fsoiltaxa <- soiltaxa %>%
  filter(
    covariate %in% c(
      "Eczema symptoms",
      "Rhinitis symptoms",
      "Wheezing & asthma symptoms",
      "Allergy categories count"
    )
  )
ggplot(fsoiltaxa, aes(y = covariate, x = name, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 5) +
  xlab("Sand/soil/dust ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))
```


# Allergy-specific bacteria in skin (radEmu)

(this code could take a while to run!)

```{r rademu-skin-samples}
y1 <- runs %>%
  add_prevalence()

ggplot(y1$taxa, aes(x = prevalence)) +
  geom_histogram(binwidth = 1) +
  xlim(c(20, 160)) +
  # scale_y_log10()+
  labs(x = "Prevalence in skin samples", y = "ASV count") +
  geom_vline(xintercept = 96.5, col = "blue")

y1$samples <- drop_na(y1$samples, allergy_symptoms)
y1$counts <- filter(y1$counts, sample_id %in% y1$samples$sample_id)

y1$taxa <- y1$taxa %>%
  arrange(desc(prevalence)) %>%
  left_join(taxn) %>%
  mutate(keep = F) %>%
  {
    .$keep[1:200] <- T
    .
  }

y1 <- y1 %>%
  mutate_taxa(name = ifelse(keep == FALSE, "other", taxon_name)) %>%
  select_taxa(taxon_id, name) %>%
  aggregate_taxa() %>%
  add_prevalence()

mostprevskin <- y1$taxa %>%
  left_join(rune$taxa) %>%
  select(name, family, class, phylum, prevalence)
#write.csv(mostprevskin, "mostprevskin.csv", quote = F, row.names = F) # for supplementary materials

samp <- select(y1$samples,
               sample_id,
               Eczema,
               Rhinitis,
               Wheezing,
               allergy_symptoms,
               school_id)
rownames(samp) <- samp$sample_id
y <- tidy_count_to_matrix(y1$counts)[samp$sample_id, ]

xfun::cache_exec(
  {
    fit1 <- radEmu::emuFit(
      formula = ~ Eczema,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(k = "Eczema", j = seq_len(ncol(y)))
    )
    fit2 <- radEmu::emuFit(
      formula = ~ Rhinitis,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(k = "Rhinitis", j = seq_len(ncol(y)))
    )
    fit3 <- radEmu::emuFit(
      formula = ~ Wheezing,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(k = "Wheezing", j = seq_len(ncol(y)))
    )
    fit4 <- radEmu::emuFit(
      formula = ~ allergy_symptoms,
      data = samp,
      Y = y,
      cluster = samp$school_id,
      test_kj = data.frame(k = "allergy_symptoms", j = seq_len(ncol(y)))
    )
  },
  path = "source/extra_microbiome_analyses_cache/"
)

ecz <- fit1$coef %>%
  filter(covariate == "Eczema") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
ecz

rhi <- fit2$coef %>%
  filter(covariate == "Rhinitis") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
rhi

whe <- fit3$coef %>%
  filter(covariate == "Wheezing") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
whe

sym <- fit4$coef %>%
  filter(covariate == "allergy_symptoms") %>%
  mutate(taxon_id = category) %>%
  left_join(y1$taxa) %>%
  select(covariate, name, estimate, pval)
sym

skintaxa <- rbind(ecz, rhi, whe, sym)

sigtax <- skintaxa %>%
  filter(pval < 0.05 &
           covariate %in% c("Eczema", "Rhinitis", "Wheezing", "allergy_symptoms"))
skintaxa <- skintaxa %>%
  filter(name %in% sigtax$name) %>%
  mutate(significance = ifelse(pval < 0.05, "*", "")) %>%
  mutate(
    covariate = fct_recode(
      covariate,
      "Eczema symptoms" = "Eczema",
      "Rhinitis symptoms" = "Rhinitis",
      "Wheezing & asthma symptoms" = "Wheezing",
      "Allergy categories count" = "allergy_symptoms",
      "Urban landscape" = "landscape_typelow naturalness",
      "Grey schoolground" = "school_contextlow greenness",
      "Natural elements score" = "natural_elements_score"
    )
  ) %>%
  mutate(
    covariate = fct_relevel(
      covariate,
      "Urban landscape",
      "Grey schoolground",
      "Natural elements score",
      "Eczema symptoms",
      "Rhinitis symptoms",
      "Wheezing & asthma symptoms",
      "Allergy categories count"
    )
  )

ggplot(skintaxa, aes(y = covariate, x = name, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 5) +
  xlab("Skin ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))
```


# Models for class diversity

```{r}
data <- data %>%
  mutate(
    loglib = log(lib_size),
    log_skin_simp = log(skin_simpson),
    log_skin_rich = log(skin_richness)
  )

kruskal.test(Gamma_richness ~ school_id, data = data)

p <- c()
est <- c()
outcome <- c()
rs <- c()
class <- c()
plots <- c()
for (c in classes) {
  cl <- classdiv %>%
    filter(classes == c)
  dat <- data %>%
    select(
      Eczema,
      Wheezing,
      Rhinitis,
      biological_sex,
      garden_score,
      log_skin_simp,
      log_skin_rich,
      rich = cl$namesrich,
      simp = cl$namessimp,
      loglib,
      landscape_type,
      school_context,
      school_id
    ) %>%
    mutate(log_rich = log(rich), log_simp = log(simp))
  lm <- glmer(
    Eczema ~ log_simp * log_skin_simp + biological_sex + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "InverseSimpson")
  outcome <- append(outcome, "Eczema")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  lm <- glmer(
    Rhinitis ~ log_simp * log_skin_simp + biological_sex + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "InverseSimpson")
  outcome <- append(outcome, "Rhinitis")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  lm <- glmer(
    Wheezing ~ log_simp * log_skin_simp + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "InverseSimpson")
  outcome <- append(outcome, "Wheezing")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  lm <- glmer(
    Eczema ~ log_rich * log_skin_rich + biological_sex + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "Richness")
  outcome <- append(outcome, "Eczema")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  lm <- glmer(
    Rhinitis ~ log_rich * log_skin_rich + biological_sex + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "Richness")
  outcome <- append(outcome, "Rhinitis")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  lm <- glmer(
    Wheezing ~ log_rich * log_skin_rich + (1 | school_id),
    family = binomial,
    data = dat
  )
  rs <- append(rs, "Richness")
  outcome <- append(outcome, "Wheezing")
  p <- append(p, coef(summary(lm))[2, 4])
  est <- append(est, coef(summary(lm))[2, 1])
  class <- append(class, c)
  kr <- kruskal.test(log_rich ~ school_context, data = dat)
  rs <- append(rs, "None")
  outcome <- append(outcome, "alogrich")
  p <- append(p, kr$p.value)
  est <- append(est, as.numeric(kr$statistic))
  class <- append(class, c)
  plots <- append(
    plots,
    ggplot(dat, aes(x = school_context, y = log_rich)) +
      geom_boxplot() +
      labs(title = c))
}
glmer_outcome <- tibble(class, outcome, rs, est, p) %>%
  filter(rs %in% c("Richness", "None"))
```


# Permanovas for skin

skin community variation significantly linked with wheezing and asthma (R2 = 0.003)
no effect from pollution
by far largest effect from school ID

```{r}
perform_adonis(runs, c("Eczema", "Wheezing", "Rhinitis", "school_id")) # wheezing
perform_adonis(runs, c("pm10", "sirm", "bc", "Eczema", "school_id"))
perform_adonis(runs, c("pm10", "sirm", "bc", "Rhinitis", "school_id"))
perform_adonis(runs, c("pm10", "sirm", "bc", "Wheezing", "school_id"))
perform_adonis(
  runs,
  c(
    "biological_sex",
    "natural_elements_score",
    "nature_exposure",
    "school_context",
    "landscape_type",
    "ses",
    "school_ses",
    "school_id"
  )
) # "ses" not signficant AFTER school

rsq <- c()
pv <- c()
factors <- c(
  "sirm",
  "t_score_phy",
  "Eczema",
  "Rhinitis",
  "Wheezing",
  "allergy_symptoms",
  "school_id",
  "ses",
  "landscape_type",
  "school_context",
  "natural_elements_score",
  "nature_connectedness"
) # "no2" , "pm10", "pm25", "bc",

for (f in factors) {
  ad <- perform_adonis(
    runs,
    c(f, "biological_sex"),
    by = "margin",
    permutations = 999
  )
  rsq <- append(rsq, ad$R2[1])
  pv <- append(pv, ad$`Pr(>F)`[1])
}

perm <- tibble(factors, rsq, pv) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "no2",
      "pm10",
      "pm25",
      "bc",
      "sirm",
      "allergy_symptoms",
      "Wheezing",
      "Rhinitis",
      "Eczema",
      "t_score_phy",
      "ses",
      "nature_connectedness",
      "school_id",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  )) # %>%
# filter(factors%!in% c("t_score_phy","no2" , "pm10", "pm25", "bc"))

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_rect(
    aes(
      xmin = as.numeric(factors) - 0.4,
      xmax = as.numeric(factors) + 0.4,
      ymin = 0.001, ymax = rsq
    ),
    fill = "darkblue"
  ) +
  # geom_col(fill="darkblue")+
  geom_point(shape = 8, size = 1, aes(y = rsq + rsq * 0.2, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  scale_y_continuous(
    trans = "log10",
    limits = c(0.001, 1), # avoid zero (log undefined) & give room
    breaks = c(0.001, 0.01, 0.05, 0.1, 0.2, 1),
    labels = scales::label_number()
  ) +
  scale_x_discrete(
    labels = c(
      "school_id" = "School ID",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "ses" = "Child's SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "Schoolyard greenness",
      "landscape_type" = "Landscape type",
      "nature_connectedness" = "Nature connectedness",
      "Wheezing" = "Wheezing and asthma",
      "t_score_phy" = "Physical activity and health",
      "allergy_symptoms" = "Allergy categories count"
    )
  ) +
  xlab("") +
  ylab("PERMANOVA R squared (beta diversity) skin") +
  # ylim(values=c(0,0.2))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(margin = margin(r = -8)),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p
```


# Permanovas for environment

The adonis2 function was modified to perform permutations only for the variable of interest and with the parameter by=’margin’.
For each effect of interest, we tested the following model: ~ e_i + landscape type + sample type.
In this model, e_i is the effect of interest, landscape type is low or high naturalness environment of the school, sample type is the origin of the microbiome (plant, sand, dirt, or dust).

```{r}
means <- runs$samples %>%
  drop_na(t_score_phy) %>%
  group_by(school_id) %>%
  summarize(mean_score_phy = mean(t_score_phy),
            ne = sum(!is.na(t_score_phy))) %>%
  mutate(mean_score_phy = ifelse(ne < 3, NA, mean_score_phy)) %>%
  select(-ne) %>%
  full_join(allergies)
rune$samples <- rune$samples %>%
  left_join(means)

# for all env samples

rsq <- c()
pv <- c()
factors <- c(
  "sample_type",
  "sirm",
  "school_id",
  "school_ses",
  "landscape_type",
  "school_context",
  "natural_elements_score"
) # , "mean_score_phy", "meaneczema", "meanrhinitis", "meanwheezing", "allergy_symptoms","no2" , "pm10", "pm25", "bc",
for (f in factors) {
  ad <- perform_adonis(
    rune,
    c(f, "landscape_type", "sample_type"),
    by = "margin",
    permutations = 999
  )
  rsq <- append(rsq, ad$R2[1])
  pv <- append(pv, ad$`Pr(>F)`[1])
}
perm <- tibble(factors, rsq, pv) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "sample_type",
      "sirm",
      "school_id",
      "school_ses",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  )) # , "allergy_symptoms", "meanwheezing", "meanrhinitis", "meaneczema", "mean_score_phy""no2" , "pm10", "pm25", "bc",

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_rect(
    aes(
      xmin = as.numeric(factors) - 0.4,
      xmax = as.numeric(factors) + 0.4,
      ymin = 0.001,
      ymax = rsq
    ),
    fill = "darkblue"
  ) +
  geom_point(shape = 8, aes(y = rsq + rsq * 0.2, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  scale_y_continuous(
    trans = "log10",
    limits = c(0.001, 0.5),
    # avoid zero (log undefined) & give room
    breaks = c(0.001, 0.01, 0.1, 0.3),
    labels = scales::label_number()
  ) +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "school_id" = "School ID",
      "sample_type" = "Sample type",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "School context",
      "landscape_type" = "Landscape type"
    )
  ) +
  xlab("") +
  ylab("PERMANOVA R squared\nall environmental samples") +
  # ylim(values=c(0,0.27))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p

# for soil/dust env samples only


rund <- rune %>%
  filter_samples(sample_type %in% c("dust", "sand", "dirt"))

rsq <- c()
pv <- c()
factors <- c(
  "sample_type",
  "no2",
  "pm10",
  "pm25",
  "bc",
  "sirm",
  "school_id",
  "school_ses",
  "landscape_type",
  "school_context",
  "natural_elements_score"
) # "mean_score_phy", "meaneczema", "meanrhinitis", "meanwheezing", "allergy_symptoms",
for (f in factors) {
  ad <- perform_adonis(rund,
                       c(f, "landscape_type"),
                       by = "margin",
                       permutations = 999)
  rsq <- append(rsq, ad$R2[1])
  pv <- append(pv, ad$`Pr(>F)`[1])
}
perm <- tibble(factors, rsq, pv) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "sample_type",
      "no2",
      "pm10",
      "pm25",
      "bc",
      "sirm",
      "school_id",
      "school_ses",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  )) # "allergy_symptoms", "meanwheezing", "meanrhinitis", "meaneczema", "mean_score_phy",
p <- ggplot(perm, aes(x = factors, y = rsq * 100)) +
  geom_col(fill = "darkblue") +
  geom_point(shape = 8, aes(y = rsq * 100 + 0.5, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "school_id" = "School ID",
      "sample_type" = "Sample type",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "School context",
      "landscape_type" = "Landscape type"
    )
  ) +
  xlab("") +
  ylab("Percent of community variation explained") +
  # ylim(values=c(0,0.47))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p

# for plant samples only

runp <- rune %>%
  filter_samples(sample_type == "plant")
rsq <- c()
pv <- c()
factors <- c(
  "sirm",
  "school_id",
  "school_ses",
  "landscape_type",
  "school_context",
  "natural_elements_score"
) # "no2" , "pm10", "pm25", "bc",

for (f in factors) {
  ad <- perform_adonis(runp,
                       c(f, "landscape_type"),
                       by = "margin",
                       permutations = 999)
  rsq <- append(rsq, ad$R2[1])
  pv <- append(pv, ad$`Pr(>F)`[1])
}

perm <- tibble(factors, rsq, pv) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "sirm",
      "school_id",
      "school_ses",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  )) # "no2" , "pm10", "pm25", "bc",

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_rect(
    aes(
      xmin = as.numeric(factors) - 0.4,
      xmax = as.numeric(factors) + 0.4,
      ymin = 0.001,
      ymax = rsq
    ),
    fill = "darkblue"
  ) +
  geom_point(shape = 8, aes(y = rsq + rsq * 0.2, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  scale_y_continuous(
    trans = "log10",
    limits = c(0.001, 1),
    # avoid zero (log undefined) & give room
    breaks = c(0.001, 0.01, 0.1, 0.5),
    labels = scales::label_number()
  ) +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "school_id" = "School ID",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "School context",
      "landscape_type" = "Landscape type"
    )
  ) +
  xlab("") +
  ylab("PERMANOVA R squared\nplant samples") +
  # ylim(values=c(0,0.27))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p



# for dust samples only

runp <- rune %>%
  filter_samples(sample_type == "dust")
rsq <- c()
pv <- c()
factors <- c(
  "no2",
  "pm10",
  "pm25",
  "bc",
  "sirm",
  "school_id",
  "school_ses",
  "landscape_type",
  "school_context",
  "natural_elements_score"
)

for (f in factors) {
  ad <- perform_adonis(runp,
                       c(f, "landscape_type"),
                       by = "margin",
                       permutations = 999)
  rsq <- append(rsq, ad$R2[1])
  pv <- append(pv, ad$`Pr(>F)`[1])
}

perm <- tibble(factors, rsq, pv) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "no2",
      "pm10",
      "pm25",
      "bc",
      "sirm",
      "school_id",
      "school_ses",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  ))

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_rect(
    aes(
      xmin = as.numeric(factors) - 0.4,
      xmax = as.numeric(factors) + 0.4,
      ymin = 0.001,
      ymax = rsq
    ),
    fill = "darkblue"
  ) +
  geom_point(shape = 8, aes(y = rsq + rsq * 0.2, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  scale_y_continuous(
    trans = "log10",
    limits = c(0.001, 1),
    # avoid zero (log undefined) & give room
    breaks = c(0.001, 0.01, 0.1, 0.5),
    labels = scales::label_number()
  ) +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "school_id" = "School ID",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "School context",
      "landscape_type" = "Landscape type"
    )
  ) +
  xlab("") +
  ylab("PERMANOVA R squared\ndust samples") +
  # ylim(values=c(0,0.27))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p
```


# alphas for environment: inverse simpson

```{r}
runer <- rune %>%
  add_total_count() %>%
  add_alpha(method = "invsimpson") %>%
  add_alpha(method = "obs")

simps <- runer$samples %>%
  group_by(sample_type, school_id) %>%
  reframe(
    mean_simpson = mean(invsimpson),
    total_count = mean(total_count),
    sample_type = sample_type
  ) %>%
  unique()

obs <- runer$samples %>%
  group_by(sample_type, school_id) %>%
  reframe(mean_richness = median(obs), sample_type = sample_type) %>%
  unique()

sdata <- rune %>%
  select_samples(
    school_id,
    sample_id,
    no2,
    pm10,
    pm25,
    bc,
    sirm,
    meaneczema,
    meanrhinitis,
    meanwheezing,
    allergy_symptoms,
    school_ses,
    landscape_type,
    school_context,
    natural_elements_score,
    tree_number
  ) %>%
  aggregate_samples()

data <- simps %>%
  left_join(obs) %>%
  left_join(select(sdata$samples, -sample_id)) |> 
  mutate(
    sample_type = factor(
      sample_type,
      levels = c("dirt", "dust", "sand", "plant"),
      labels = c("ground", "ground", "ground", "plant")
    )
  )

# expect negative binomial or poisson distribution

rsq <- c()
pv <- c()
ste <- c()
factors <- c(
  "sirm",
  "school_ses",
  "landscape_type",
  "school_context",
  "natural_elements_score",
  "sample_type"
) # "allergy_symptoms", "mean_score_phy", "meaneczema", "meanrhinitis", "meanwheezing","no2" , "pm10", "pm25", "bc",
mod <- glmmTMB(
  mean_simpson ~ sirm 
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"), data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_simpson ~ school_ses
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"), data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_simpson ~ landscape_type
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"), data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_simpson ~ school_context
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"),
  data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_simpson ~ natural_elements_score
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"),
  data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_simpson ~ sample_type
  + offset(log(total_count))
  + (1 | school_id),
  family = Gamma(link = "log"),
  data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

perm <- tibble(factors, rsq, pv, ste) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(pv = round(pv, digits = 3)) %>%
  mutate(pv = ifelse(pv == 0, "< 0.001", pv)) %>%
  mutate(factors = fct_relevel(
    factors,
    c(
      "sample_type",
      "sirm",
      "school_ses",
      "natural_elements_score",
      "school_context",
      "landscape_type"
    )
  )) # "allergy_symptoms", "meanwheezing", "meanrhinitis", "meaneczema", "mean_score_phy",,"no2" , "pm10", "pm25", "bc"

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_hline(
    yintercept = 0,
    linetype = "dotted",
    color = "grey",
    size = 0.5
  ) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = rsq - ste, ymax = rsq + ste),
                width = .2,
                position = position_dodge(.9)) +
  geom_text(size = 2.5, aes(label = paste("p =", pv, sep = " "), y = rsq + ste + 0.35)) +
  geom_point(shape = 8, aes(y = rsq + ste + 0.7, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  geom_hline(yintercept = 0,
             linetype = "dotted",
             linewidth = 0.5) +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "sample_type" = "Sample type (plant)",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "Non-green schoolyard",
      "landscape_type" = "Landscape type (urban)"
    )
  ) +
  xlab("") +
  ylab("Log-scale effect on Inverse Simpson\n(including all environmental samples)") +
  #  ylim(values=c(0,0.02))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p
```


# alphas for environment: richness

```{r}
data$mean_richness <- as.integer(data$mean_richness)
ggplot(data, aes(mean_richness)) +
  geom_histogram()

rsq <- c()
pv <- c()
ste <- c()
factors <- c("sirm", "school_ses", "landscape_type", "school_context", "natural_elements_score", "sample_type") # "no2" , "pm10", "pm25", "bc",

mod <- glmmTMB(
  mean_richness ~ sirm
  + offset(log(total_count))
  + (1 | school_id),
  family = poisson, data = data)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_richness ~
    school_ses 
  + offset(log(total_count))
  + (1 | school_id),
  family = poisson, data = data)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_richness ~ landscape_type
  + offset(log(total_count))
  + (1 | school_id),
 family = poisson, data = data)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_richness ~ school_context 
  + offset(log(total_count))
  + (1 | school_id),
  family = poisson,
  data = data
)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_richness ~ natural_elements_score
  + offset(log(total_count))
  + (1 | school_id),
  family = poisson, data = data)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

mod <- glmmTMB(
  mean_richness ~ sample_type
  + offset(log(total_count))
  + (1 | school_id),
 family = poisson, data = data)
rsq <- append(rsq, coef(summary(mod))$cond[2, 1])
pv <- append(pv, coef(summary(mod))$cond[2, 4])
ste <- append(ste, coef(summary(mod))$cond[2, 2])

perm <- tibble(factors, rsq, pv, ste) %>%
  mutate(significant = ifelse(pv < 0.05, "T", "F")) %>%
  mutate(pv = round(pv, digits = 3)) %>%
  mutate(pv = ifelse(pv == 0, "< 0.001", pv)) %>%
  mutate(factors = fct_relevel(factors, c("sample_type", "sirm", "school_ses", "natural_elements_score", "school_context", "landscape_type"))) # "allergy_symptoms", "meanwheezing", "meanrhinitis", "meaneczema", "mean_score_phy",,"no2" , "pm10", "pm25", "bc",

p <- ggplot(perm, aes(x = factors, y = rsq)) +
  geom_hline(
    yintercept = 0,
    linetype = "dotted",
    color = "grey",
    size = 0.5
  ) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = rsq - ste, ymax = rsq + ste),
                width = .2,
                position = position_dodge(.9)) +
  geom_text(size = 2.5, aes(label = paste("p =", pv, sep = " "), y = rsq + ste + 0.3)) +
  geom_point(shape = 8, aes(y = rsq + ste + 0.6, alpha = significant)) +
  scale_alpha_manual(values = c(0, 1), guide = "none") +
  geom_hline(yintercept = 0,
             linetype = "dotted",
             size = 0.5) +
  #  scale_y_continuous(trans="log10") +
  scale_x_discrete(
    labels = c(
      "sample_type" = "Sample type (plant)",
      "no2" = "Modelled NO2",
      "pm10" = "Modelled PM10",
      "pm25" = "Modelled PM2.5",
      "bc" = "Modelled black carbon",
      "sirm" = "SIRM (local air pollution)",
      "school_ses" = "School mean SES",
      "natural_elements_score" = "Natural elements score",
      "school_context" = "Non-green schoolyard",
      "landscape_type" = "Urban landscape type"
    )
  ) +
  xlab("") +
  ylab("Log-scale effect on ASV richness\n(including all environmental samples)") +
  #  ylim(values=c(0,0.02))+
  coord_flip() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    panel.background = element_rect(fill = "transparent"),
    plot.background = element_rect(fill = "transparent", color = NA)
  )
p
```


# Transferred ASVs

```{r}
planttaxa <- planttaxa %>%
  rename(taxon_name = name) %>%
  left_join(taxn)

runp <- rune %>%
  #  filter_samples(sample_type=="plant")%>%
  filter_counts(count > 3) %>%
  mutate_samples(sample_type_gen = "environmental") %>%
  #  filter_taxa(taxon%in%c(planttaxa$taxon))%>% # only plant ASVs related to allergies
  select_samples(sample_id, sample_type_gen, sample_type, school_id)
schools <- unique(runp$samples$school_id)

runm <- runs %>%
  filter_counts(count > 3) %>%
  select_samples(
    sample_id,
    child_id,
    school_id,
    Eczema,
    Wheezing,
    Rhinitis,
    landscape_type,
    school_context,
    natural_elements_score
  ) %>%
  mutate_samples(sample_type = "skin") %>%
  mutate_samples(sample_type_gen = "skin") %>%
  filter_samples(school_id %in% schools) %>%
  merge_tidytacos(runp, taxon_identifier = taxon) %>%
  select_taxa(-species) %>%
  add_taxon_name()

taxov <- taxn %>%
  rename(taxon_name_overall = taxon_name) %>%
  left_join(select(runm$taxa, taxon_name, taxon))
# this code generates a table with, for each skin taxon, at how many schools there was an overlap between plant/environment and skin

zeros <- (rep(0, nrow(runm$taxa)))
taxnames <- select(runm$taxa, taxon, taxon_name)
tax <- unique(taxnames$taxon_name)
taxon_name <- tax
overlap <- tibble(taxon_name, number = zeros)

for (sid in schools) {
  envtax <- runp %>%
    filter_samples(school_id == sid) %>%
    taxa()
  sktax <- runs %>%
    filter_samples(school_id == sid) %>%
    taxa()
  taxon <- intersect(sktax$taxon, envtax$taxon)
  overl <- tibble(taxon) %>%
    left_join(taxnames, by = join_by(taxon))
  for (t in tax) {
    if (t %in% overl$taxon_name) {
      overlap <- overlap %>%
        mutate(number = ifelse(taxon_name == t, number + 1, number))
    }
  }
}

occ <- runm %>%
  add_prevalence(condition = "sample_type_gen") %>%
  taxa() %>%
  select(taxon_name, prevalence = prevalence_in_environmental)
overlap <- overlap %>%
  filter(number > 0) %>%
  left_join(taxov)
overlapf <- overlap %>%
  filter(number > 10)
# overlapf<-overlapf%>%
#  mutate(type="transfer")

occ <- occ %>%
  filter(taxon_name %in% overlapf$taxon_name)

# this code assesses how often plant/env taxa occurred on cheeks

transfer <- runs %>%
  filter_counts(count > 3) %>%
  filter_taxa(taxon %in% runp$taxa$taxon) %>% # only environmental ASVs
  add_prevalence(relative = T) %>%
  taxa() %>%
  left_join(taxov) %>%
  select(taxon_name_overall, child = prevalence)

trov <- overlap %>%
  mutate(school = number / 30) %>%
  left_join(transfer, by = join_by(taxon_name_overall)) %>%
  select(-taxon_name, -taxon, -number)

trov2 <- trov %>%
  pivot_longer(!taxon_name_overall, names_to = "type_of_transfer", values_to = "number")
```



```{r fig-transfer-plant-child-1, fig.height=200/25.4}
# this figure shows transfer per child in the plant ASV heatmap
planttaxa2 <- planttaxa %>%
  #  filter(pval<0.05)%>%
  rename(taxon_name_overall = taxon_name) %>%
  select(-taxon)
transfer2 <- transfer %>%
  filter(taxon_name_overall %in% planttaxa2$taxon_name_overall) %>%
  mutate(
    covariate = "Transferred",
    pval = NA,
    significance = round(child, digits = 2)
  ) %>% # text in block
  mutate(estimate = -child * 20) %>% # color adjustement
  select(-child)
planttaxa2 <- rbind(planttaxa2, transfer2)

ggplot(planttaxa2,
       aes(y = covariate, x = taxon_name_overall, fill = estimate)) +
  geom_tile() +
  scale_fill_gradient2(low = "darkgreen", mid = "white", high = "darkorange") +
  geom_text(aes(label = significance), color = "black", size = 2) +
  xlab("Plant ASVs") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))
```


```{r fig-transfer-plant-child-2}
ggplot(transfer2, aes(x = taxon_name_overall, y = significance)) +
  geom_bar(stat = "identity") +
  ylab("prev. in children") +
  xlab("") +
  theme_bw() +
  theme(axis.text.x = element_blank())
```




