---
title: "Supplementary information for _Environmental bacteria link green schoolyards to better child well-being and fewer allergy symptoms_"
author: "Hans Van Calster, Wenke Smets, Marie Legein, Wannes Van Beeck, Lize Delanghe, Antoine Groslambert, Harmony Brulein, Els Lommelen, Linda Van Meersche, Michael Leone, Edith Swerts, Arlien Wuyts, Karen Wuyts, Max Dekeukeleire, Liesa Lammens, Melike Özen, Rafiqa Benchrih, Mina Diri, Saskia Wanner, Sofie Heyman, Raïsa Carmen, Roeland Samson, Eva M. De Clercq, Anna Leonard, Sophie O. Vanwambeke, Raf Aerts, Hans Keune, Sarah Lebeer, Irina Spacova"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
bibliography: health_paper.yaml
---

```{r setup, include=FALSE}
# follow the instructions in file source/reproduce.R
library(knitr)
library(here)
opts_chunk$set(
    echo = TRUE, message = FALSE, out.width = "100%",
  dev = "png", dpi = 300)
opts_knit$set(root.dir = here())
library(frictionless)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(modelsummary)
library(performance)
library(patchwork)
library(brms)
library(tidybayes)
library(bayestestR)
library(INBOtheme)
library(bayesplot)
library(posterior)
library(kableExtra)

conflicted::conflicts_prefer(brms::ar)
conflicted::conflicts_prefer(tidyr::expand)
conflicted::conflicts_prefer(brms::lognormal)
conflicted::conflicts_prefer(brms::loo)
conflicted::conflicts_prefer(tidyr::pack)
conflicted::conflicts_prefer(tidyr::unpack)
conflicted::conflicts_prefer(dplyr::filter)
conflicted::conflicts_prefer(dplyr::lag)

theme_set(theme_inbo(transparent = TRUE))
model_folder <- here(
    "source", "model_objects_paper"
  )
data_objects_folder <- here(
    "source", "data_objects_paper"
)

fs::dir_create(model_folder)
fs::dir_create(data_objects_folder)
```

```{r cache=TRUE, results='hide'}
1 + 1
# just a trick to prevent the folder containing figures from being removed
# when knitting a self-contained document
```

```{css, echo=FALSE}
img {
  max-width:150%;
  height: auto;
}

.warning {
  background-color: #fcf8e3;
  color: #8a6d3b;
  padding: 15px;
  border-left: 5px solid #faebcc;
  border-radius: 4px;
  margin-bottom: 15px;
}
```

```{r colours}
school_colours <- c(
  `green` = "#40B0A6",
  `non-green` = "#E1BE6A"
)
landscape_colours <- c(rural = "cadetblue", urban = "sienna")
```

# Data preparation

::: warning
**Warning:** This document contains the analysis pipeline that was used for the paper *Environmental bacteria link green schoolyards to better child well-being and fewer allergy symptoms*.
While the code is identical, the data used in this document (513 participants) is a (large) subset of the data used in the paper (529 participants).
The reason for this is that some participants / parents did not give consent for long-term storage / archiving of their data.
As such, we had permission to use their data in publications, but not to publish their data in a long-term public archive.
The data used in this repository, which are downloaded from <https://doi.org/10.5281/zenodo.10527033> (see `source/reproduce.R`) [@vancalster2024], therefore only include data from participants / parents that gave consent for long-term storage of their data.
As a consequence, the results from the paper are not exactly reproducible.
:::

```{r cleaned-data}
cleaned_data <- read_package(
  here::here(
    "data",
    "datapackage_an_zenodo_cleaned_data",
    "datapackage.json"
  )
)
```

```{r derived-data}
derived_data <- read_package(
  here::here(
    "data",
    "datapackage_an_zenodo_derived_data",
    "datapackage.json"
  )
)

# resources(derived_data)
dd_school_level <- read_resource(
  derived_data, "wp1_school_level_key_variables"
) |>
  mutate(
    landscape_type = factor(
      landscape_type, levels = c("low naturalness", "high naturalness")),
    school_context = factor(
      school_context, levels = c("low greenness", "high greenness")
    )
  )

dd_child_level <- read_resource(
  derived_data, "wp1_child_level_key_variables"
)


kidscreen_groups <- read_resource(
  cleaned_data,
  resource_name = "wp4_kidscreen_questions"
) |>
  distinct(question_group) |>
  rename(label = question_group) |>
  mutate(
    level = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )

# reorder
kidscreen_groups <- kidscreen_groups[c(5,2,1,3,4),]

kidscreen_groups$label[kidscreen_groups$level == "tscorephy"] <-
  "Physical well-being"

kidscreen_groups$short_label <- c(
  "At school", "Psychological", "Physical", "At home", "Social"
)
```

```{r health-data}
health_data <- dd_child_level |>
  group_by(school_id) |>
  mutate(
    ses_mean = mean(ses, na.rm  = TRUE),
    ses_centered = ses - ses_mean,
  ) |>
  select(-ses, -ses_dimensions, -ses_dimensions_relative) |>
  left_join(
    dd_school_level |>
      select(
        school_id,
        community,
        province_name,
        landscape_type,
        school_context,
        case_control_id,
        ses_school = ses,
        natural_elements_score,
        tree_number,
        artificial_elements_score,
        allergenic_potential,
        mean_log_sirm,
        mean_sirm_logratio,
        pm25
      ),
    by = join_by(school_id)) |>
  mutate(
    eczema_01 = ifelse(eczema < 0, 0, 1),
    rhinitis_01 = ifelse(rhinitis < 0, 0, 1),
    wheezing_01 = ifelse(wheezing < 0, 0, 1),
    n_sensitive = eczema_01 + rhinitis_01 + wheezing_01
  ) |>
  ungroup()
```

```{r calc-nature-exposure}
outdoor_play <- read_resource(
  package = cleaned_data,
  resource_name = "wp5_parents_outdoor_play_data_part2"
)
outdoor_play <- outdoor_play |>
    filter(
      question_code %in% 
        c("x5_5", "x5_7", "x5_8", "x5_9", "x5_11", "x5_12", "x5_13", "x5_14",
          "x5_15", "x5_17", "x5_18")) |>
    mutate(response = ifelse(
      response == "no association - association",
      0.5,
      ifelse(response == "association",
             1,
             ifelse(response == "no association",
                    0,
                    NA)
             )
      )
    ) |>
  summarise(
    response = mean(response, na.rm = TRUE),
    .by = c(child_id, question_code)
  ) |>
  summarise(
    nature_exposure = sum(response) / 11,
    .by = c(child_id)
  )

health_data <- health_data |>
  left_join(outdoor_play, by = "child_id")
```

```{r garden-data, warning=FALSE}
living_env <- read_resource(
  cleaned_data, "wp5_parents_living_environment_data")

garden_data_recoded <- living_env |>
  filter(grepl("^garden", question_short)) |>
  mutate(
    response_recoded = case_when(
      question_short == "garden_present" & response_values == "1" ~ 0,
      question_short == "garden_present" & response_values != "1" ~ 1,
      question_short == "garden_size" & response_values == "1" ~ 0,
      question_short == "garden_size" & response_values == "2" ~ 1,
      question_short == "garden_size" & response_values == "2 - 3" ~ 1.5,
      question_short == "garden_size" & response_values == "3" ~ 2,
      question_short == "garden_size" & response_values == "4" ~ 3,
      question_short == "garden_type" & response_values == "1" ~ 0,
      question_short == "garden_type" & response_values == "1 - 3" ~ 1,
      question_short == "garden_type" & response_values == "2" ~ 1,
      question_short == "garden_type" & response_values == "3" ~ 2,
      question_short == "garden_type" & response_values == "4" ~ 1.5,
      question_short == "garden_usage" ~ as.numeric(response_values) - 1
    )
  )

garden_data <- garden_data_recoded |>
  pivot_wider(
    id_cols = c(child_id, replicate_id, case),
    names_from = question_short,
    values_from = response_recoded,
    values_fill = list(response_recoded = 0)) |>
  mutate(
    garden_score = (garden_size / 3 + garden_type / 2 + garden_usage / 3) / 3
  ) |>
  select(child_id, garden_present, garden_score) |>
  group_by(child_id) |> 
  summarise(
    garden_present = max(garden_present, na.rm = TRUE),
    garden_score = mean(garden_score, na.rm = TRUE)
  ) |>
  ungroup()

health_data <- health_data |>
  left_join(garden_data, by = "child_id")
```

```{r add-allergy-family-data}
allergy_related <- read_resource(
  cleaned_data, "wp5_parents_allergy_related_questions_data")

recode_table <- readr::read_csv2(
     here("data", "recode_table_allergy_questions.csv")
   )

allergy_related <- allergy_related |>
  inner_join(recode_table,
             by = join_by(question_short, response_values, interpretation)) |>
  select(child_id, question_short, recode_value)

# some are not uniquely defined, using max to collapse them
allergy_related <- allergy_related |>
  pivot_wider(
    names_from = question_short,
    values_from = recode_value,
    values_fn = max
  )

allergy_related <- allergy_related |>
  mutate(
    sum_contact_animals = rowSums(across(starts_with("contact")))
    ) |>
  select(!(starts_with("contact") | disease_hinders_atop))

health_data <- health_data |>
  left_join(
    allergy_related,
    by = join_by(child_id)
  ) |> 
  mutate(
    f_allergies_family = case_when(
      is.na(allergies_family) ~ "unknown",
      allergies_family == 0 ~ "no",
      allergies_family == 1 ~ "yes"
    )
  )

```

```{r read-microbiome-diversity}
env_samples_diversity <- readr::read_csv(
  here("data", "env_samples_diversity.csv"),
  show_col_types = FALSE
)
skin_samples_diversity <- readr::read_csv(
  here("data", "skin_samples_diversity.csv"),
  show_col_types = FALSE
)
```

```{r add-diversity}
health_data <- health_data |>
  left_join(
    skin_samples_diversity,
    by = join_by(child_id)
  ) |>
  left_join(
    env_samples_diversity,
    by = join_by(school_id)
  ) |> 
  mutate(
    log_inverse_simpson_skin = log(inverse_simpson_skin),
    log_invsimpson_plant = log(invsimpson_plant),
    log_invsimpson_dust = log(invsimpson_dust),
    log_invsimpson_soil = log(invsimpson_soil),
    log_total_count_skin = log(total_count_skin)
  )
```

```{r rescale-t-scores}
# rescale t-scores (which by construction have a mean of 50 and sd of 10)
health_data <- health_data |>
  mutate(
    across(
      starts_with("t_score"),
      \(x) (x - 50) / 10)
  )
```

```{r add-missing-natural-elements-score}
# some schools have no natural elements score
# these missing values were imputed by manual interpretation of photo's 
# from the playgrounds
health_data <- health_data |>
  mutate(
    natural_elements_score = case_when(
      school_id == "01" ~ 1 / 16,
      school_id == "11" ~ (1 + 1 + 1 + 1 + (2 / 75) ^ 0.2) / 16,
      school_id == "27" ~ 3 / 16,
      school_id == "32" ~ mean(
        c(
          (1 + 1 + 1 + 1 + (15 / 75) ^ 0.2) / 16,
          (1 + 1 + 1 + 1 + (20 / 75) ^ 0.2) / 16,
          (1 + 1) / 16 
        )),
      school_id == "36" ~ (1 + 1 + 1 + 1 + 1 + 1 + (10 / 75) ^ 0.2) / 16,
      TRUE ~ natural_elements_score
    ))
```

# Data analysis

## School and participant characteristics

A total of 167 schools were contacted out of which 40 expressed interest to participate and 37 participated; these 37 schools had 929 children in the fifth grade out of which 529 children (57%) gave consent to participate in at least one of the tests (Fig. \@ref(fig:flowcharts)) [@aerts2025].
In addition, 16 children did not give consent for long-term archiving of data.

(ref:flowcharts) Flowchart depicting the number of schools that were contacted, interested and participated (left-side, and the number of fifth grade children that gave consent and for which at least one test was available (right-side).

```{r flowcharts, fig.cap="(ref:flowcharts)", out.width=c("100%")}
knitr::include_graphics(
  here("media", "flowchart.png"),
  rel_path = FALSE
)
```

Low naturalness landscape schools were typically found in urbanized areas and characterized by high levels of impervious surfaces and low levels of combined forest and grassland cover.
High naturalness landscape schools were found in rural areas with low levels of impervious surfaces and high levels of combined forest and grassland cover.
We will henceforth refer to these as schools in urban or rural landscapes.

```{r relabel}
health_data <- health_data |>
  mutate(
    school_context = factor(
      school_context,
      levels = c("low greenness", "high greenness"),
      labels = c("non-green", "green")
      ),
    landscape_type = factor(
      landscape_type,
      levels = c("low naturalness", "high naturalness"),
      labels = c("urban", "rural")
      )
  )
```

## Descriptive statistics of the study population

Table \@ref(tab:table1) shows descriptive statistics of the main variables in our sample.
The number of participating children in schools labelled as 'non-green' is about 1.5 times higher compared to those in 'green' schools.

(ref:table1) Description of main characteristics of the study population. *N* is the number of children.

```{r table1}
data_to_summarize <- health_data |>
  select(
    landscape_type, school_context,
    ses_mean, ses_centered, biological_sex,
    birth_year, community,
    clusters = school_id
  ) |>
  mutate(
    community = ifelse(
      community == "Dutch speaking.French speaking",
      "French speaking", community),
    lt_sc = interaction(landscape_type, school_context, lex.order = TRUE)
  )

t0 <- modelsummary::datasummary_balance(
  formula = ses_mean + ses_centered + birth_year + biological_sex +
    community ~ lt_sc,
  data = data_to_summarize,
  dinm = FALSE,
  title = "(ref:table1)",
  output = "kableExtra"
  )

t0
```

## Exploration health outcomes

Missingness is about 4% for ISAAC data and 10% for KIDSCREEN data.

```{r}
health_outcomes <- health_data |>
  select(
    starts_with("t_score"),
    ends_with("_01")
  )

health_outcomes |>
  naniar::vis_miss() +
  theme(axis.text.x = element_text(vjust = 0))
```

We calculate Spearman rank correlations between outcomes and visualise them in Figure \@ref(fig:corrplot).

The KIDSCREEN scores are all positively correlated with each other.

- phy: physical well-being
- sch: well being at school
- par: family life and free time
- pwb: general mood and how you feel
- soc: friends

The ISAAC scores are also correlated but to a lesser extent.

(ref:corrplot) Spearman rank correlation matrix among modelled outcome variables.

```{r corrplot, fig.cap="(ref:corrplot)"}
cordat <- health_outcomes |>
  cor(method = "spearman", use = "pairwise.complete.obs")

corrplot::corrplot(cordat)
```

## Exploration of covariates

### Missingness

```{r potential-covariates}
covars <- ~
  landscape_type +
  school_context +
  biological_sex +
  ses_mean +
  ses_centered +
  pm25 +
  natural_elements_score +
  nature_connectedness +
  garden_score +
  mean_log_sirm +
  inverse_simpson_skin +
  invsimpson_dust +
  invsimpson_plant +
  invsimpson_soil +
  sum_contact_animals +
  allergies_family
```

Missingness patterns among covariates (Figure \@ref(fig:missing-covars)).

(ref:missing-covars) Variables (columns) are sorted in order of amount of missingness and observations (rows) are clustered according to missingness.

```{r missing-covars, fig.cap = "(ref:missing-covars)"}
health_data |>
  select(
    all_of(all.vars(covars))) %>%
  mutate(all_combined = ifelse(complete.cases(.), 1, NA)) |>
  naniar::vis_miss(cluster = TRUE, sort_miss = TRUE) +
  expand_limits(x = 22) +
  theme(axis.text.x = element_text(vjust = 0))
```

### Collinearity

We checked Spearman rank correlations between covariates (Figure \@ref(fig:collinearity1)).

The strongest correlation between continuous covariates equals 0.52 and is between `pm25` and `ses_mean`.
This is a modest correlation and it is likely not problematic to include both covariates in a model.

(ref:collinearity1) Correlation plot showing Spearman rank correlation between covariates.

```{r collinearity1, fig.cap = "(ref:collinearity1)", fig.height = 6, fig.width = 6}
cordat2 <- health_data |>
  select(
    all_of(all.vars(covars))) |>
  select(where(is.double)) |>
  cor(method = "spearman", use = "pairwise.complete.obs")

corrplot::corrplot(
  cordat2,
  order = "AOE",
  diag = FALSE,
  addCoef.col = 'darkgrey',
  number.cex = 0.8)
```

We also look at relationships between categorical and continuous covariates (Figure \@ref(fig:collinearity2)).
The resulting relationships do not indicate collinearity problems.
One minor concern is that the range of `ses_mean` values within high naturalness landscape for high greenness schools is more restricted compared to the other factor combinations.

(ref:collinearity2) Distribution of covariate values as a function of landscape type, school context and biological sex.

```{r collinearity2, fig.height=200/25.4, fig.cap = "(ref:collinearity2)"}
health_data |>
  select(
    all_of(all.vars(covars))) |>
  pivot_longer(
    cols = where(is.double),
    names_to = "covariate",
    values_to = "value"
  ) |>
  mutate(
    context = recode_values(
      covariate,
        "ses_mean" ~ "school",
        "ses_centered" ~ "home",
        "pm25" ~ "school",
        "natural_elements_score" ~ "school",
        "nature_connectedness" ~ "child",
        "garden_score" ~ "home",
        "mean_log_sirm" ~ "school",
        "inverse_simpson_skin" ~ "child",
        "invsimpson_dust" ~ "school",
        "invsimpson_plant" ~ "school",
        "invsimpson_soil" ~ "school",
        "sum_contact_animals" ~ "home",
        "allergies_family"   ~ "home"
    )
  ) |>
  filter(!is.na(value), !is.na(biological_sex)) |>
  arrange(context, covariate) |> 
  mutate(facetvar = paste0(context, " - ", covariate)) |> 
  ggplot() +
  stat_sum(
    aes(
      x = landscape_type,
      y = value,
      colour = school_context,
      shape = biological_sex
    ),
    alpha = 0.5,
    position = position_dodge(width = 0.5)
  ) +
  scale_colour_manual(values = school_colours) +
  facet_wrap(
    ~ facetvar |> str_replace_all("_", " ") |> str_wrap(15),
    scales = "free_y") +
  theme(legend.position = "top") +
  guides(
    colour = guide_legend(ncol =  1, direction = "vertical"),
    shape = guide_legend(ncol = 2, direction = "vertical"),
    size = guide_legend(nrow = 1, direction = "vertical")
  ) +
  scale_x_discrete(labels = \(x) str_wrap(x, 10))
```

## KIDSCREEN-27 model

We model only all five kidscreen dimensions.
Included all two-way interactions between mean SES, Schoolyard and landscape type.
The scores are modelled with a Gaussian distribution and an identity link.
All missing data are imputed during model fit.
Missing data are declared as parameters to be estimated by there own sets of regression equations that include the interaction between schoolyard and landscape type, as well as a random intercept for school in case of child-level varying covariates.
For the main equations, correlations between the school random intercepts are modelled as well.

```{r adj-mi-ks, warning=FALSE}
rhs_ks <- ~ mi(ses_mean) +
  landscape_type +
  school_context +
  landscape_type:school_context +
  mi(ses_mean):school_context +
  mi(ses_mean):landscape_type +
  #child-level characteristics
  biological_sex +
  mi(ses_centered) +
  mi(garden_score) +
  mi(nature_connectedness) +
  #school-level characteristics
  natural_elements_score +
  #scale(pm25) +
  mi(mean_log_sirm) +
  (1 | q | school_id)

bf_adj_mi_ks <- 
  bf(
    update(rhs_ks, t_score_phy | mi() ~ .),
    decomp = "QR",
    family = gaussian()
  ) +
    bf(
    update(rhs_ks, t_score_par | mi() ~ .),
    decomp = "QR",
    family = gaussian()
  ) +
  bf(
    update(rhs_ks, t_score_pwb | mi() ~ .),
    decomp = "QR",
    family = gaussian()
  ) +
  bf(
    update(rhs_ks, t_score_sch | mi() ~ .),
    decomp = "QR",
    family = gaussian()
  ) +
  bf(
    update(rhs_ks, t_score_soc | mi() ~ .),
    decomp = "QR",
    family = gaussian()
  ) +
  bf(
    nature_connectedness | mi() ~ school_context*landscape_type + (1|school_id),
    family = gaussian()
  ) +
  bf(
    ses_centered | mi() ~ school_context*landscape_type + (1|school_id),
    family = gaussian()
  ) +
  bf(
    ses_mean | mi() + trunc(lb = 0, ub = 1) ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    garden_score | mi() ~ school_context*landscape_type + (1|school_id),
    family = gaussian()
  ) +
  bf(
    mean_log_sirm | mi() + trunc(lb = 0) ~ school_context*landscape_type,
    family = gaussian()
  ) +
  set_rescor(FALSE)

priors <- set_prior(
  prior = "normal(0, 2)", class = "b",
  resp = c("gardenscore", "meanlogsirm", "natureconnectedness", 
"sescentered", "sesmean", "tscorephy")
  )


mv_adj_mi_ks <- brm(
    formula = bf_adj_mi_ks,
    data = health_data,
    prior = priors,
    cores = 4,
    silent = 2,
    refresh = 0,
    chains = 4,
    backend = "cmdstanr",
    normalize = FALSE,
    file_refit = "on_change",
    file = file.path(
      model_folder,
      "mv_adj_mi_ks"
    )
)
```

```{r summary-ks, warning=FALSE}
pvals <- pd(mv_adj_mi_ks, verbose = FALSE) |>
  as_tibble() |>
  separate_wider_delim(
    Parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE)) |>
  select( -Effects, -Component)

gather_draws(mv_adj_mi_ks, `b_tsc.*|bsp_tsc.*`, regex = TRUE) |>
  separate_wider_delim(
    .variable,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  group_by(response, parameter) |>
  summarise_draws() |>
  inner_join(pvals, by = join_by(response, parameter)) |>
  relocate(p_value, .before = rhat) |>
  select(-variable, -rhat, -ess_bulk, -ess_tail) |>
  kable(digits = 2)
```

```{r mcmc-intervals-ks}
# Calculate centering and scaling constants
mean_ses <- mean(health_data$ses_mean, na.rm = TRUE)
sd_ses_mean <- stats::sd(health_data$ses_mean, na.rm = TRUE)
sd_nature_connectedness <- stats::sd(
  health_data$nature_connectedness,
  na.rm = TRUE
)
sd_ses_centered <- stats::sd(health_data$ses_centered, na.rm = TRUE)
sd_garden_score <- stats::sd(health_data$garden_score, na.rm = TRUE)
sd_natural_elements_score <- stats::sd(
  health_data$natural_elements_score,
  na.rm = TRUE
)
sd_mean_log_sirm <- stats::sd(health_data$mean_log_sirm, na.rm = TRUE)

# Extract the draws object from the brms model
draws_df <- as_draws_df(mv_adj_mi_ks)

responses <- c("tscorephy", "tscorepar", "tscorepwb", "tscoresch", "tscoresoc")

for (resp in responses) {
  # ---------------------------------------------------------
  # RE-CENTER CATEGORICAL MAIN EFFECTS (Evaluated at Mean SES)
  # ---------------------------------------------------------

  draws_df[[paste0("b_", resp, "_school_contextgreen")]] <-
    draws_df[[paste0("b_", resp, "_school_contextgreen")]] + 
    (draws_df[[paste0("bsp_", resp, "_mises_mean:school_contextgreen")]] * mean_ses)
  
  draws_df[[paste0("b_", resp, "_landscape_typerural")]] <-
    draws_df[[paste0("b_", resp, "_landscape_typerural")]] + 
    (draws_df[[paste0("bsp_", resp, "_mises_mean:landscape_typerural")]] * mean_ses)

  # ---------------------------------------------------------
  # SCALE CONTINUOUS COVARIATES (To represent a 1-SD change)
  # ---------------------------------------------------------
  
  # SES Mean (bsp_)
  draws_df[[paste0("bsp_", resp, "_mises_mean")]] <- 
    draws_df[[paste0("bsp_", resp, "_mises_mean")]] * sd_ses_mean
    
  draws_df[[paste0("bsp_", resp, "_mises_mean:school_contextgreen")]] <- 
    draws_df[[paste0("bsp_", resp, "_mises_mean:school_contextgreen")]] * sd_ses_mean
    
  draws_df[[paste0("bsp_", resp, "_mises_mean:landscape_typerural")]] <- 
    draws_df[[paste0("bsp_", resp, "_mises_mean:landscape_typerural")]] * sd_ses_mean
  
  # Other continuous covariates
  draws_df[[paste0("bsp_", resp, "_minature_connectedness")]] <- 
    draws_df[[paste0("bsp_", resp, "_minature_connectedness")]] * sd_nature_connectedness
    
  draws_df[[paste0("bsp_", resp, "_mises_centered")]] <- 
    draws_df[[paste0("bsp_", resp, "_mises_centered")]] * sd_ses_centered
    
  draws_df[[paste0("bsp_", resp, "_migarden_score")]] <- 
    draws_df[[paste0("bsp_", resp, "_migarden_score")]] * sd_garden_score
    
  draws_df[[paste0("bsp_", resp, "_mimean_log_sirm")]] <- 
    draws_df[[paste0("bsp_", resp, "_mimean_log_sirm")]] * sd_mean_log_sirm
}

pvals <- pd(draws_df) %>%
  as_tibble() %>%
  filter(grepl("^(b_|bsp_)", x = Parameter)) |> 
  separate_wider_delim(
    Parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) %>%
  select(-dropme) %>%
  mutate(p_value = pd_to_p(pd))

ks_alldata <- mcmc_intervals_data(
  draws_df,
  regex_pars = "^(b|bsp)_tscore",
  prob = 0.5, # inner interval probability mass
  prob_outer = 0.9, # outer interval probability mass
  point_est = "median"
) |>
  separate_wider_delim(
    parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) %>%
  select(-dropme) %>%
  inner_join(
    pvals,
    by = c("response", "parameter")
  ) |> 
  mutate(
    parameter = gsub("landscape_typerural", "Rural", parameter),
    parameter = gsub("school_contextgreen", "Green schoolyard", parameter),
    parameter = gsub("biological_sexFemale", "Biological sex - girl", parameter),
    parameter = gsub("natural_elements_score", "Natural elements score", parameter),
    parameter = gsub("mises_centered", "Deviation school SES", parameter),
    parameter = gsub("mises_mean", "Mean school SES", parameter),
    parameter = gsub("migarden_score", "Garden score", parameter),
    parameter = gsub("mimean_log_sirm", "Local air pollution", parameter),
    parameter = gsub("minature_connectedness", "Nature connectedness", parameter)
  )
```

Figure \@ref(fig:mcmc-intervals-kidscreen-all-data)

(ref:mcmc-intervals-kidscreen-all-data) Parameter estimates for the effects on physical well-being. Point estimates represent the median of the posterior distribution. Thick and thin lines show 50% and 90% inner probability mass of the posterior distribution. Effects are with reference to urban landscape type, boys and non-green schoolyards. Categorical effects were evaluated at the mean of the 'Mean school SES' variable. Continuous covariates were scaled so that estimates represent the change associated with a one standard deviation increase.

```{r mcmc-intervals-kidscreen-all-data, fig.height=6, out.width="100%", fig.cap= "(ref:mcmc-intervals-kidscreen-all-data)", message=FALSE}
mcmc_intervals_kidscreen_all_data <- ks_alldata |>
  mutate(
    `Well-being` = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    ),
    parameter = reorder(parameter, desc(m))
  ) |> 
  filter(
    parameter != "Intercept") |>
  mutate(
    p_val_label = case_when(
      p_value <= 0.05 ~ "*",
      p_value <= 0.1 ~ "^",
      TRUE ~ ""
    )
  ) |> 
  ggplot() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_linerange(
    aes(
      x = m * 10,
      xmin = ll * 10,
      xmax = hh * 10,
      y = parameter,
      colour = `Well-being`
    ),
    linewidth = 1, alpha = 0.5, position = position_dodge(width = 0.7)
  ) +
  geom_linerange(
    aes(
      x = m * 10,
      xmin = l * 10,
      xmax = h * 10,
      y = parameter,
      colour = `Well-being`
    ),
    linewidth = 2, alpha = 0.5, position = position_dodge(width = 0.7)
  ) +
  geom_point(
    aes(
      x = m * 10,
      y = parameter,
      colour = `Well-being`
    ),
    size = 3, alpha = 0.5, position = position_dodge(width = 0.7)
  ) +
  geom_text(
    aes(
      x = hh * 10 + 1,
      y = parameter,
      label = p_val_label,
      colour = `Well-being`
    ),
    position = position_dodge(width = 0.7), show.legend = FALSE
  ) +
  labs(
    x = "Estimated effect on self-reported well-being (KIDSCREEN-27)"
  ) +
  theme(axis.title.y = element_blank())

mcmc_intervals_kidscreen_all_data

saveRDS(
  mcmc_intervals_kidscreen_all_data,
  file = file.path(data_objects_folder, "mcmc_intervals_kidscreen_all_data.rds")
)
```

```{r ppcheck-ks, warning=FALSE}
responses <- mv_adj_mi_ks$formula$responses
checks <- vector("list", length(responses))
for (i in seq_along(responses)) {
  checks[[i]] <- pp_check(
    mv_adj_mi_ks,
    resp = responses[i],
    ndraws = 50) + labs(x = responses[i])
}
checks |>
  wrap_plots()
```

We can calculate post-hoc comparison, stratified by landscape type to see if school types differ significantly within landscape types:

```{r emmeans-sclt-ks}
refgrid1 <- emmeans::ref_grid(
  mv_adj_mi_ks,
  at = list(
    ses_mean = median(mv_adj_mi_ks$data$ses_mean, na.rm = TRUE),
    landscape_type = c("urban", "rural"),
    rep.meas = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )
)

emm <- emmeans::emmeans(
  refgrid1,
  specs = ~ school_context | ses_mean + rep.meas + landscape_type)

contr <- emmeans::contrast(
  emm,
  method = "revpairwise",
  # multiple by 10 (difference cancels out the offset of 50) to obtain original scale 
  scale = 10
  )

pvals <- pd(contr) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE))

pe_ks_contrasts_sc <- confint(contr) |> 
  as_tibble() |>
  inner_join(pvals) |> 
  rename(response = rep.meas) |> 
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
      )
  )
pe_ks_contrasts_sc |> 
  kable(digits = 2)
```

Figure \@ref(fig:adj-mi-ks-preds-sclt-school-effects) shows the interaction between schoolyard and landscape-type with and without including the random effects of each school for the five KIDSCREEN indicators.

(ref:adj-mi-ks-preds-sclt-school-effects) KIDSCREEN-27 score predictions (median and 95% credible intervals) as a function of landscape type and schoolyard. The fixed effects are shown (larger symbols, no transparancy) as well as the effects including the random effects for each school (smaller symbols, transparant). The p-values are for the differences between green and non-green schoolyards within a landscape type.

```{r adj-mi-ks-preds-sclt-school-effects, out.width="120%", fig.cap = "(ref:adj-mi-ks-preds-sclt-school-effects)"}
preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = "landscape_type:school_context",
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
)


preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ks_sc_lt <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = "school_id",
  conditions = make_conditions(
    mv_adj_mi_ks,
    c("landscape_type", "school_context")
  ),
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch"),
  re_formula = NULL
  )


preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

school_cats <- health_data |>
  distinct(school_id, landscape_type, school_context)

preds_df <- preds_df |>
  semi_join(school_cats, by = join_by(school_id, landscape_type, school_context))

preds_df_ks_sc_lt_school_effects <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

pe_ks_contrasts_sc <- pe_ks_contrasts_sc |> 
  mutate(
    star = case_when(
      p_value <= 0.05 ~ "*",
      p_value <= 0.1 ~ "^",
      TRUE ~ ""
    ),
    label = sprintf("p = %.02f%s", p_value, star)
  )

p_ks_sc_lt_school_effects <- preds_df_ks_sc_lt_school_effects |>
    # Group by the facet and x-axis variable
  group_by(response, landscape_type) |>
  mutate(
    sign = mean(estimate__[school_context == "green"]) -
      mean(estimate__[school_context == "non-green"]),
    sortorder = sign * estimate__
  ) |> 
  # Sort by sign * estimate__
  arrange(sortorder) |> 
  # Create a local rank for the dodging order
  mutate(
    dodge_rank = factor(row_number())
  ) |> 
  ungroup() |> 
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      colour = school_context,
      group = dodge_rank
    ),
    linewidth = 0.2,
    size = 0.2,
    alpha = 0.4,
    position = position_dodge(width = 0.8)
  ) +
  geom_pointrange(
    data = preds_df_ks_sc_lt,
    aes(
      x = landscape_type,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      colour = school_context
    ),
    position = position_dodge(width = 0.8)
  ) +
  geom_text(
    data = pe_ks_contrasts_sc |> arrange(response, landscape_type),
    aes(
      x = landscape_type,
      y = c(60, 55, 60, 55, 55, 50, 60, 50, 60, 55),
      label = label
    )
  ) +
  scale_colour_manual(values = school_colours) +
  scale_y_continuous(breaks = seq(30, 60, 5)) +
  facet_grid(cols = vars(response), labeller = label_wrap_gen(width = 15)) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "KIDSCREEN-27 score\n(+/- 95% CI)"
  )

saveRDS(
  p_ks_sc_lt_school_effects,
  file = file.path(data_objects_folder, "p_ks_sc_lt_school_effects.rds")
)
ggsave(
  plot = p_ks_sc_lt_school_effects,
  filename = file.path(data_objects_folder, "p_ks_sc_lt_school_effects.png"),
  width = 30, height = 15, units = "cm", bg = "white",
  dpi = 320
)
knitr::include_graphics(
  file.path(data_objects_folder, "p_ks_sc_lt_school_effects.png"),
  rel_path = FALSE
)
```

Figure \@ref(fig:sei-sc-interaction-ks) shows the interaction between school-level SES, on the one hand, and Schoolyard or landscape type, on the other hand.

(ref:sei-sc-interaction-ks) Predictions for the two-way interaction effects between school-level socio-economic index and Schoolyard, and school-level socio-economic index and landscape type, on physical well-being (backtransformed to original scale). The regression line depicts the median of the posterior distribution and the shaded areas correspond with 95% credible intervals.

```{r sei-sc-interaction-ks, fig.height=140/25.4, out.width="130%", fig.cap="(ref:sei-sc-interaction-ks)"}
preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = "ses_mean:school_context",
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )

preds_df_ses_sc <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

# restrict to observed ranges of ses
ses_ranges <- health_data |>
  group_by(landscape_type, school_context) |>
  summarise(
    min_ses_mean = min(ses_mean, na.rm = TRUE),
    max_ses_mean = max(ses_mean, na.rm = TRUE)
) |>
  ungroup()

saveRDS(
  ses_ranges,
  file.path(data_objects_folder, "ses_ranges.rds")
)

preds_df_ses_sc <- preds_df_ses_sc |>
  inner_join(ses_ranges) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_sc <- preds_df_ses_sc |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

p_ks_ses_sc <- preds_df_ses_sc |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      colour = school_context
    )) +
  scale_y_continuous(breaks = seq(30, 60, 5)) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_grid(cols = vars(response), labeller = label_wrap_gen(width = 15)) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "KIDSCREEN-27 score +/- 95% CI"
  )

preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = "ses_mean:landscape_type",
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )

preds_df_ses_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_lt <- preds_df_ses_lt |>
  inner_join(ses_ranges) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_lt <- preds_df_ses_lt |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()


p_ks_ses_lt <- preds_df_ses_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      fill = landscape_type
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      colour = landscape_type
    )) +
  scale_y_continuous(breaks = seq(30, 60, 5)) +
  scale_colour_manual(
    values = landscape_colours) +
  scale_fill_manual(
    values = landscape_colours) +
  facet_grid(cols = vars(response), labeller = label_wrap_gen(width = 15)) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Landscape type",
    fill = "Landscape type",
    y = "KIDSCREEN-27 score +/- 95% CI"
  )

p_ks_sc_lt <-  ggplot() +
  geom_pointrange(
    data = preds_df_ks_sc_lt,
    aes(
      x = landscape_type,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      colour = school_context
    ),
    position = position_dodge(width = 0.8)
  ) +
  scale_colour_manual(values = school_colours) +
  scale_y_continuous(breaks = seq(30, 60, 5)) +
  facet_grid(cols = vars(response), labeller = label_wrap_gen(width = 15)) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "KIDSCREEN-27 score +/- 95% CI"
  )

p_ks_sc_lt / p_ks_ses_sc / p_ks_ses_lt +
  plot_layout(guides = "collect", axes = "collect", axis_titles = "collect") +
  plot_annotation(tag_levels = "a")

saveRDS(
  p_ks_ses_lt,
  file.path(data_objects_folder, "p_ks_ses_lt.rds")
)
saveRDS(
  p_ks_ses_sc,
  file.path(data_objects_folder, "p_ks_ses_sc.rds")
)
```

Alternative visualisation that better shows how the three two-way interactions combine:

Figure \@ref(fig:sei-sc-interaction-ks-2) shows the two-way interactions between school-level SES, schoolyard and landscape type.

(ref:sei-sc-interaction-ks-2) Predictions for the two-way interaction effects between school-level socio-economic index, Schoolyard and landscape type, on KIDSCREEN-27 outcomes (backtransformed to original scale). The regression line depicts the median of the posterior distribution and the shaded areas correspond with 95% credible intervals.

```{r sei-sc-interaction-ks-2, fig.height=200/25.4, out.width="100%", fig.cap="(ref:sei-sc-interaction-ks-2)"}
preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = "ses_mean:school_context",
  conditions = make_conditions(mv_adj_mi_ks, "landscape_type"),
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )

preds_df_ses_sc_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_sc_lt <- preds_df_ses_sc_lt |>
  inner_join(ses_ranges) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  ) |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

p_ks_ses_sc_lt <- preds_df_ses_sc_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__ * 10 + 50,
      colour = school_context
    )) +
  scale_y_continuous(breaks = seq(30, 60, 5)) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_grid(response ~ landscape_type) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "KIDSCREEN-27 score +/- 95% CI"
  )

p_ks_ses_sc_lt

saveRDS(
  p_ks_ses_sc_lt,
  file.path(data_objects_folder, "p_ks_ses_sc_lt.rds")
)

```

```{r}
# Obtain slopes for each landscape, school 
emt <- emmeans::emtrends(
  mv_adj_mi_ks,
  specs = c( "school_context", "landscape_type", "rep.meas"),
  var = "ses_mean",
  rep.meas = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
)

pvals_emt <- pd(emt) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE))

annot <- print(emt, export = TRUE)
```

Table \@ref(tab:slopes-2way) shows the corresponding slopes.

(ref:slopes-2way) `r paste(annot$annotations, collapse = ". ")`. The slopes are scaled such that a 0.1 unit increase in SEI corresponds with an increase of the KIDSCREEN-27 score equal to the slope value. For instance, a 0.1 unit increase in SEI, increases the "At school" score by 3.00 [CI: 0.85 - 5.26] in non-green, urban schools.

```{r slopes-2way}
confint(emt) |> 
  as_tibble() |>
  inner_join(pvals_emt) |>
  filter(grepl("^tscore", rep.meas)) |>
  mutate(
    response = factor(
      rep.meas,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )) |> 
   select(-rep.meas) |> 
    kable(digits = 2, caption = "(ref:slopes-2way)")
```

We can further calculate post-hoc differences at specific SEI values allowing to check for a remediation effect.

```{r}
refgrid1 <- emmeans::ref_grid(
  mv_adj_mi_ks,
  at = list(
    ses_mean = seq(0.4, 0.7, 0.1),
    landscape_type = c("urban", "rural"),
    rep.meas = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
  )
)


emm <- emmeans::emmeans(
  refgrid1,
  specs = ~ school_context | ses_mean + rep.meas + landscape_type)

contr <- emmeans::contrast(
  emm,
  method = "revpairwise",
  # differences multiply by 10 to have original scale (offset of 50 cancels)
  scale = 10
  )


pvals <- pd(contr) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE))

annot <- print(contr, export = TRUE)
```

Table \@ref(tab:emmeans-ses-phy) shows contrasts between green and non-green schoolyard for specified levels of SES and landscape type.

(ref:emmeans-ses-phy) `r paste(annot$annotations, collapse = ". ")`. pd = probability of direction, p_value = 2 \* (1 - pd). The estimates show the difference between green and non-green schoolyards, e.g. in urban context, for a SEI of 0.4, children score 5.17 [CI: 0.20 - 10.43] points higher for physical well-being when schools have green schoolyards compared to non-green schoolyards. This significant effect disappears for higher SEI values in the urban context which is indicative of a remediating effect of green schoolyards partially offsetting negative effects of low SEI in urban settings.

```{r emmeans-ses-phy}
confint(contr) |> 
  as_tibble() |>
  inner_join(pvals) |> 
    mutate(
    response = factor(
      rep.meas,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )) |> 
  select(-rep.meas) |> 
  kable(digits = 2, caption = "(ref:emmeans-ses-phy)")
```

```{r conditional-effects-adjusters-ks}
adjusters <- c(
  "biological_sex",
  "ses_centered",
  "garden_score",
  "natural_elements_score",
  "nature_connectedness",
  "mean_log_sirm")


preds <- conditional_effects(
  mv_adj_mi_ks,
  effects = adjusters,
  resp = c("tscorephy", "tscorepwb", "tscorepar", "tscoresoc", "tscoresch")
)

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    adjuster <- names(x)[1]
    x$response <- response
    x$effect1__ <- adjuster
    x
  }
)

preds_df <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()
```

```{r}
p_ks_adjusters <- preds_df |>
  select(-biological_sex, -landscape_type, -school_context) |>
  pivot_longer(
    cols = any_of(adjusters),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(
    variable == effect1__
  ) |>
  mutate(effect1__ = gsub("_", " ",effect1__)) |>
  ggplot() +
  geom_ribbon(
    aes(x = value,
        y = estimate__ * 10 + 50,
        ymin = lower__ * 10 + 50,
        ymax = upper__ * 10 + 50),
    alpha = 0.4
    ) +
  geom_line(
    aes(
      x = value,
      y = estimate__ * 10 + 50
    )
  ) +
  facet_grid(
    response~stringr::str_wrap(effect1__, 15),
    scales = "free") +
  theme(
    legend.position = "top",
    legend.box = "vertical",
    legend.margin = margin()
  ) +
  labs(y = "KIDSCREEN-27 score +/- 95% CI") 

p_ks_adjusters
saveRDS(
  p_ks_adjusters,
  file.path(data_objects_folder, "p_ks_adjusters.rds")
)

```

```{r ks-biological-sex}
p_ks_biological_sex <- preds_df |>
  filter(effect1__ == "biological_sex") |>
  ggplot() +
  geom_pointrange(
    aes(
      x = response,
      y = estimate__ * 10 + 50,
      ymin = lower__ * 10 + 50,
      ymax = upper__ * 10 + 50,
      colour = biological_sex
    ),
    position = position_dodge(width = 0.2)
  ) +
  labs(y = "KIDSCREEN-27 score +/- 95% CI") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

p_ks_biological_sex
saveRDS(
  p_ks_biological_sex,
  file.path(data_objects_folder, "p_ks_biological_sex.rds")
)
```

## Wheezing, rhinitis, eczema (ISAAC survey data)

```{r adj-mi-isaac-new, warning=FALSE}
priors <- set_prior(
  prior = "normal(0,1)", class = "b", 
  resp = c(
    "eczema01", "rhinitis01", "wheezing01",
  "sesmean", "loginversesimpsonskin")
) 

rhs_isaac <- ~
  #designvars
  landscape_type +
  school_context +
  landscape_type:school_context +
  mi(ses_mean) +
  mi(ses_mean):school_context +
  mi(ses_mean):landscape_type +
  mi(log_inverse_simpson_skin) +
  mi(log_inverse_simpson_skin):school_context +
  mi(log_inverse_simpson_skin):landscape_type +
  biological_sex +
  f_allergies_family +
  (1|q|school_id)

bf_adj_mi_isaac_ses_skin <-
  bf(
    update(rhs_isaac, eczema_01 ~ .),
    decomp = "QR",
    family = bernoulli()
  ) +
  bf(
    update(rhs_isaac, rhinitis_01 ~ . + mi(allergenic_potential)),
    decomp = "QR",
    family = bernoulli()
  ) +
  bf(
    update(rhs_isaac, wheezing_01 ~ .  + mi(allergenic_potential)),
    decomp = "QR",
    family = bernoulli()
  ) +
  bf(
    ses_mean | mi() + trunc(lb = 0, ub = 1) ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    log_inverse_simpson_skin | mi() ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    allergenic_potential | mi() + trunc(lb = 0) ~ school_context*landscape_type,
    family = gaussian()
  ) +
  set_rescor(FALSE)

mv_adj_mi_isaac_ses_skin <- brm(
    formula = bf_adj_mi_isaac_ses_skin,
    data = health_data,
    prior = priors,
    cores = 4,
    silent = 2,
    refresh = 0,
    chains = 4,
    backend = "cmdstanr",
    normalize = FALSE,
    file_refit = "on_change",
    file = file.path(
      model_folder,
      "mv_adj_mi_isaac_ses_skin")
)

# 22 rows containing either missing response variable
# or missing biological sex are excluded
# because stan cannot impute discrete parameters
# n remaining equals 491
```

```{r summary-isaac-new}
pvals <- pd(mv_adj_mi_isaac_ses_skin) |>
  as_tibble() |>
  separate_wider_delim(
    Parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE)) |>
  select(-Effects, -Component)

gather_draws(
  mv_adj_mi_isaac_ses_skin,
  `b_(ecz|rhi|whe).*|bsp_(ecz|rhi|whe).*`,
  regex = TRUE
) |>
  separate_wider_delim(
    .variable,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  group_by(response, parameter) |>
  summarise_draws() |>
  inner_join(pvals) |>
  relocate(p_value, .before = rhat) |>
  select(-variable, -rhat, -ess_bulk, -ess_tail) |>
  mutate(
    estimate = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      q5,
      q95
    )
  ) |> 
  select(response, parameter, estimate, p_value) |>
  kable(digits = 2)
```

\@ref(fig:mcmc-intervals-isaac-new)

(ref:mcmc-intervals-isaac-new) Parameter estimates for the effects on eczema, rhinitis and wheezing (binary outcomes). Point estimates represent the median of the posterior distribution. Thick and thin lines show 50% and 90% inner probability mass of the posterior distribution. For the all data model, effects are with reference to urban landscape type, boys and non-green schoolyards. For the rural and uban subsets, effects are with reference to boys and non-green schoolyards.

```{r mcmc-intervals-isaac-new, fig.height = 200/25.4, out.width="120%", fig.cap="(ref:mcmc-intervals-isaac-new)"}
isaac_alldata <- mcmc_intervals_data(
  mv_adj_mi_isaac_ses_skin,
  regex_pars = "^(b|bsp)_(eczema01|rhinitis01|wheezing01)",
  prob = 0.5, # inner interval probability mass
  prob_outer = 0.9, # outer interval probability mass
  point_est = "median"
) |> 
  mutate(
    dataset = "all data"
  )

isaac_alldata <- isaac_alldata |> 
  mutate(
    response = stringr::str_extract(string = parameter, "^(b|bsp)_(.+)01_", group = 2),
    parameter = gsub("^(b|bsp)_(.+)01_", "", parameter),
    parameter = gsub("landscape_typerural", "Rural", parameter),
    parameter = gsub("school_contextgreen", "Green schoolyard", parameter),
    parameter = gsub("biological_sexFemale", "Biological sex - girl", parameter),
    parameter = gsub("natural_elements_score", "Natural elements score", parameter),
    parameter = gsub("mises_centered", "Deviation school SES", parameter),
    parameter = gsub("mises_mean", "Mean school SES", parameter),
    parameter = gsub("migarden_score", "Garden score", parameter),
    parameter = gsub("mimean_log_sirm", "Local air pollution", parameter),
    parameter = gsub("minature_connectedness", "Nature connectedness", parameter),
    parameter = gsub("milog_inverse_simpson_skin", "Diversity skin microbiome", parameter),
    parameter = gsub("misum_contact_animals", "Contact animals", parameter),
    parameter = gsub("f_allergies_familyyes", "Allergies family - yes", parameter),
    parameter = gsub("f_allergies_familyunknown", "Allergies family - unknown", parameter),
    parameter = gsub("miallergenic_potential", "Allergenic potential", parameter),
    parameter = gsub("milog_invsimpson_plant", "Diversity plant microbiome", parameter),
    parameter = gsub("milog_invsimpson_dust", "Diversity dust microbiome", parameter),
    parameter = gsub("milog_invsimpson_soil", "Diversity soil microbiome", parameter),
    parameter = reorder(parameter, m)
  )

mcmc_intervals_isaac <- isaac_alldata |>
  filter(
    parameter != "Intercept"
  ) |> 
  ggplot() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_linerange(
    aes(
      x = m,
      xmin = ll,
      xmax = hh,
      y = parameter
    ),
    linewidth = 1, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_linerange(
    aes(
      x = m,
      xmin = l,
      xmax = h,
      y = parameter
    ),
    linewidth = 2, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_point(
    aes(
      x = m,
      y = parameter
    ),
    size = 3, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  facet_wrap(~ response, scales = "free_x") +
  labs(
    x = "Parameter estimates"
  ) +
  theme(
    legend.position = "top",
    axis.title.y = element_blank()
  )

mcmc_intervals_isaac
saveRDS(
  mcmc_intervals_isaac,
  file.path(data_objects_folder, "mcmc_intervals_isaac.rds")
)
```

Posterior predictive checks:

```{r ppcheck-isaac-new, warning=FALSE}
responses <- mv_adj_mi_isaac_ses_skin$formula$responses
checks <- vector("list", length(responses))
for (i in seq_along(responses)) {
  checks[[i]] <- pp_check(
    mv_adj_mi_isaac_ses_skin,
    resp = responses[i],
    ndraws = 50) + labs(x = responses[i])
}
checks |>
  wrap_plots()
```

Figure \@ref(fig:adj-mi-isaac-preds-sclt-new)

(ref:adj-mi-isaac-preds-sclt-new) Predicted percentage of children sensitized to rhinitis, wheezing or eczema as a function of landscape type and schoolyard. The centrality measure is the median and the line intervals correspond to 95% credible intervals.

```{r adj-mi-isaac-preds-sclt-new, fig.cap = "(ref:adj-mi-isaac-preds-sclt-new)"}
preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "landscape_type:school_context",
  resp = c("eczema01", "rhinitis01", "wheezing01")
  )

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_isaac_ltsc <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_ltsc <- preds_df_isaac_ltsc |>
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context
    ),
    position = position_dodge(width = 0.5)
  ) +
  scale_colour_manual(values = school_colours) +
  facet_wrap(~ response) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Percentage of sensitized children\n(+/- 95% CI)"
  ) +
  scale_y_continuous(labels = scales::percent)

p_isaac_ltsc

saveRDS(
  p_isaac_ltsc,
  file.path(data_objects_folder, "p_isaac_ltsc.rds")
)
```

Figure \@ref(fig:adj-mi-isaac-preds-sclt-school-effects)

(ref:adj-mi-isaac-preds-sclt-school-effects) Predicted percentage of children sensitized to rhinitis, wheezing or eczema as a function of landscape type and schoolyard. The centrality measure is the median and the line intervals correspond to 95% credible intervals.

```{r adj-mi-isaac-preds-sclt-school-effects, fig.cap = "(ref:adj-mi-isaac-preds-sclt-school-effects)"}
preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "school_id",
  conditions = make_conditions(
    mv_adj_mi_isaac_ses_skin, vars = c("landscape_type", "school_context")
  ),
  resp = c("eczema01", "rhinitis01", "wheezing01"),
  re_formula = NULL
  )

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

# filter out the actual types to which a school belongs
school_cats <- health_data |>
  distinct(school_id, landscape_type, school_context)
preds_df <- preds_df |>
  semi_join(
    school_cats,
    by = join_by(school_id, landscape_type, school_context)
  )


preds_df_isaac_ltsc_schooleffects <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_ltsc_schooleffects <- preds_df_isaac_ltsc_schooleffects |>
  # Group by the facet and x-axis variable
  group_by(response, landscape_type) |> 
  # Sort ascending by estimate__ (per your request)
  arrange(desc(estimate__)) |> 
  # Create a local rank for the dodging order
  mutate(dodge_rank = factor(row_number())) |> 
  ungroup() |> 
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context,
      # Map the group aesthetic to the local rank
      group = dodge_rank 
    ),
    size = 0.25,
    alpha = 0.5,
    position = position_dodge(width = 0.8)
  ) +
  scale_colour_manual(values = school_colours) +
  facet_wrap(~ response) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Percentage of sensitized children\n(+/- 95% CI)"
  ) +
  scale_y_continuous(labels = scales::percent)

p_isaac_ltsc_schooleffects

saveRDS(
  p_isaac_ltsc_schooleffects,
  file.path(data_objects_folder, "p_isaac_ltsc_schooleffects.rds")
)
```

We can calculate post-hoc comparison, stratified by landscape type to see if school types differ significantly within landscape types:

```{r emmeans-sclt-isaac-new}
ses_ranges_landscape <- ses_ranges |>
  group_by(landscape_type) |>
  summarise(
    min_overlap_ses_mean = max(min_ses_mean),
    max_overlap_ses_mean = min(max_ses_mean)
  )
#brms' emmeans support for multivariate models is limited to cases where all univariate models have the same family.
# instead calc manually
nd <- expand.grid(
  school_context = c("non-green", "green"),
  ses_mean = median(mv_adj_mi_isaac_ses_skin$data$ses_mean, na.rm = TRUE),
  landscape_type = c("urban", "rural")
) |>
  mutate(
    biological_sex = "Male",
    f_allergies_family = "no",
    ses_centered = 0,
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |>
      round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE)|> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE)|> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    log_inverse_simpson_skin = mean(health_data$log_inverse_simpson_skin, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    allergenic_potential = mean(
      health_data$allergenic_potential, na.rm = TRUE
    ) |>
      round(2)
  ) |>
  inner_join(ses_ranges_landscape, by = join_by(landscape_type)) |>
  filter(
    ses_mean >= min_overlap_ses_mean,
    ses_mean <= max_overlap_ses_mean
  )

# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_isaac_ses_skin,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("rhinitis01", "wheezing01", "eczema01")
)

# calculate risk ratio
pe <- pe |>
  select(
    landscape_type, school_context, .linpred
  ) |> 
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "riskratio"
    ),
    contrast = paste(last(school_context), " / ", first(school_context)),
    .by = c(landscape_type)
  )

pe_isaac_contrasts_sc <- pe |> 
  select(landscape_type, contrast, riskratio) |>
   mutate(
    rr_eczema = c(riskratio[, "eczema01"]),
    rr_rhinitis = c(riskratio[, "rhinitis01"]),
    rr_wheezing = c(riskratio[, "wheezing01"])
  ) |>
  select(-riskratio) |> 
  pivot_longer(
    cols = starts_with("rr_"),
    names_to = "response",
    names_prefix = "rr_",
    values_to = "riskratio"
  ) |>
  mutate(
    pd = pmax(sum(riskratio > 1), sum(riskratio < 1)) / sum(riskratio > -Inf),
    p_value_one_sided = pd_to_p(pd, direction = "one-sided"),
    p_value_two_sided = pd_to_p(pd, direction = "two-sided"),
    riskratio = sprintf(
      "%.2f [%.2f - %.2f]",
      median(riskratio),
      quantile2(riskratio, 0.025),
      quantile2(riskratio, 0.975)
    )
  ) |> 
  select(-pd) |> 
  mutate(
    response = factor(response, levels = c("rhinitis", "wheezing", "eczema"))
  ) |>
  arrange(response) 
pe_isaac_contrasts_sc |> 
  kable(digits = 2)

```

Figure \@ref(fig:sei-sc-interaction-isaac-new) shows the two-way interactions between school context and landscape type and between school-level SES, on the one hand, and school context or landscape type, on the other hand.

(ref:sei-sc-interaction-isaac) Two-way interaction conditional effects between school context and landscape type, school-level socio-economic index and school context, school-level socio-economic index and landscape type, on eczema, rhinitis and wheezing (backtransformed to original scale). The median of the posterior distribution and 95% credible intervals are shown.

```{r sei-sc-interaction-isaac-new, fig.height=200/25.4, out.width="120%", fig.cap="(ref:sei-sc-interaction-isaac-new)"}
preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "ses_mean:school_context",
  resp = c("eczema01", "rhinitis01", "wheezing01")
  )

preds_df_ses_sc <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_sc <- preds_df_ses_sc |>
  inner_join(ses_ranges) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_sc <- preds_df_ses_sc |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_ses_sc <- preds_df_ses_sc |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = school_context
    )) +
  facet_wrap(~response) +
  scale_y_continuous(limits = c(0, NA), labels = scales::percent) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "School context",
    fill = "School context",
    y = "Percentage of children sensitized (+/- 95% CI)"
  )

preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "ses_mean:landscape_type",
  resp = c("eczema01", "rhinitis01", "wheezing01")
  )

preds_df_ses_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_lt <- preds_df_ses_lt |>
  inner_join(ses_ranges) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_lt <- preds_df_ses_lt |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_ses_lt <- preds_df_ses_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = landscape_type
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = landscape_type
    )) +
  facet_wrap(~response) +
  scale_y_continuous(limits = c(0, NA), labels = scales::percent) +
  scale_colour_manual(
    values = landscape_colours) +
  scale_fill_manual(
    values = landscape_colours) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Landscape type",
    fill = "Landscape type",
    y = "Percentage of children sensitized (+/- 95% CI)"
  )

p_isaac_ltsc / p_isaac_ses_sc / p_isaac_ses_lt +
  plot_layout(guides = "collect", axes = "collect") +
  plot_annotation(tag_levels = "a")

saveRDS(
  p_isaac_ses_sc,
  file.path(data_objects_folder, "p_isaac_ses_sc.rds")
)
saveRDS(
  p_isaac_ses_lt,
  file.path(data_objects_folder, "p_isaac_ses_lt.rds")
)
```

Alternative visualisation that better shows how the three two-way interactions combine:

```{r isaac-sei-lt-sc-new, fig.height = 160/25.4, out.width="100%"}
preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "ses_mean:school_context",
  conditions = make_conditions(mv_adj_mi_isaac_ses_skin, "landscape_type"),
  resp = c("eczema01", "rhinitis01", "wheezing01")
  )

preds_df_ses_sc_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_sc_lt <- preds_df_ses_sc_lt |>
  inner_join(ses_ranges, by = join_by(school_context, landscape_type)) |>
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  ) |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_ses_sc_lt <- preds_df_ses_sc_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = school_context
    )) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_grid(response ~ landscape_type) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "School context",
    fill = "Schoolyard",
    y = "Percentage of children sensitized (+/- 95% CI)"
  )

p_isaac_ses_sc_lt

saveRDS(
  p_isaac_ses_sc_lt,
  file.path(data_objects_folder, "p_isaac_ses_sc_lt.rds")
)
```

The corresponding slopes are:

```{r isaac-slopes-2way-new}
# Obtain slopes for each landscape, school 
drw <- as_draws_rvars(mv_adj_mi_isaac_ses_skin)
slope_drw <- expand_grid(
  response = c("eczema", "rhinitis", "wheezing"),
  landscape_type = c("urban", "rural"),
  school_context = c("non-green", "green")
) |> 
  mutate(
    slope = case_when(
      landscape_type == "rural" & school_context == "green" & response == "eczema" ~
        drw$bsp_eczema01_mises_mean 
      + drw$`bsp_eczema01_mises_mean:landscape_typerural`
      + drw$`bsp_eczema01_mises_mean:school_contextgreen`,
      landscape_type == "rural" & school_context == "green" & response == "rhinitis" ~
        drw$bsp_rhinitis01_mises_mean 
      + drw$`bsp_rhinitis01_mises_mean:landscape_typerural`
      + drw$`bsp_rhinitis01_mises_mean:school_contextgreen`,
      landscape_type == "rural" & school_context == "green" & response == "wheezing" ~
        drw$bsp_wheezing01_mises_mean 
      + drw$`bsp_wheezing01_mises_mean:landscape_typerural`
      + drw$`bsp_wheezing01_mises_mean:school_contextgreen`,
      landscape_type == "rural" & school_context == "non-green" & response == "eczema" ~
        drw$bsp_eczema01_mises_mean 
      + drw$`bsp_eczema01_mises_mean:landscape_typerural`,
      landscape_type == "rural" & school_context == "non-green" & response == "rhinitis" ~
        drw$bsp_rhinitis01_mises_mean 
      + drw$`bsp_rhinitis01_mises_mean:landscape_typerural`,
      landscape_type == "rural" & school_context == "non-green" & response == "wheezing" ~
        drw$bsp_wheezing01_mises_mean 
      + drw$`bsp_wheezing01_mises_mean:landscape_typerural`,
      landscape_type == "urban" & school_context == "green" & response == "eczema" ~
        drw$bsp_eczema01_mises_mean 
      + drw$`bsp_eczema01_mises_mean:school_contextgreen`,
      landscape_type == "urban" & school_context == "green" & response == "rhinitis" ~
        drw$bsp_rhinitis01_mises_mean 
      + drw$`bsp_rhinitis01_mises_mean:school_contextgreen`,
      landscape_type == "urban" & school_context == "green" & response == "wheezing" ~
        drw$bsp_wheezing01_mises_mean 
      + drw$`bsp_wheezing01_mises_mean:school_contextgreen`,
      landscape_type == "urban" & school_context == "non-green" & response == "eczema" ~
        drw$bsp_eczema01_mises_mean,
      landscape_type == "urban" & school_context == "non-green" & response == "rhinitis" ~
        drw$bsp_rhinitis01_mises_mean,
      landscape_type == "urban" & school_context == "non-green" & response == "wheezing" ~
        drw$bsp_wheezing01_mises_mean
    )
  ) |> 
  rowwise() |> 
  mutate(
    median = median(slope),
    lower2.5 = quantile(slope, probs = 0.025),
    upper97.5 = quantile(slope, probs = 0.975),
    pd = pmax(
      sum(slope > 0),
      sum(slope < 0)
      ) / sum(slope > -Inf),
    p_value = pd_to_p(pd, verbose = FALSE)
  )

slope_drw |> 
  mutate(
    slope = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    ),
    response = factor(response, levels = c("rhinitis", "wheezing", "eczema"))
  ) |> 
  select(response, landscape_type, school_context, slope, p_value) |> 
  arrange(response) |> 
  kable(digits = 2)
```

```{r contrasts-remediation-isaac-new}
ses_ranges_landscape <- ses_ranges |>
  group_by(landscape_type) |>
  summarise(
    min_overlap_ses_mean = max(min_ses_mean),
    max_overlap_ses_mean = min(max_ses_mean)
  )


# semi-manual check if emmeans (next chunk) does what it needs to do
nd <- expand.grid(
  school_context = c("non-green", "green"),
  ses_mean = seq(0.4, 0.8, 0.02),
  landscape_type = c("urban", "rural")
) |>
  mutate(
    biological_sex = "Male",
    f_allergies_family = "no",
    ses_centered = 0,
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |> round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE) |> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE) |> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    log_inverse_simpson_skin = mean(health_data$log_inverse_simpson_skin, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    allergenic_potential = mean(health_data$allergenic_potential, na.rm = TRUE) |> round(2)
  ) |>
  inner_join(ses_ranges_landscape) |>
  filter(
    ses_mean >= min_overlap_ses_mean,
    ses_mean <= max_overlap_ses_mean
  )

# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_isaac_ses_skin,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("eczema01", "rhinitis01", "wheezing01")
)

# calculate logit difference and take exp which returns odds-ratio
rvar_plogis <- posterior::rfun(plogis)
pe <- pe |>
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "oddsratio"
    ),
    across(
      starts_with(".linpred"),
      \(x) rvar_plogis(last(x)) - rvar_plogis(first(x)),
      .names = "probability_difference"
    ),
    contrast = paste(last(school_context), " - ", first(school_context)),
    .by = c(ses_mean, landscape_type, biological_sex, ses_centered,
            garden_score, natural_elements_score, nature_connectedness,
            mean_log_sirm, log_invsimpson_plant,
            log_invsimpson_dust,
            log_invsimpson_soil,
            log_inverse_simpson_skin,
            sum_contact_animals,
            f_allergies_family,
            allergenic_potential)
  )

pe_pvals <- 
  pe |> 
  select(landscape_type, contrast, ses_mean,
         starts_with("oddsratio")) |> 
  mutate(
    pd = pmax(
      sum(oddsratio > 1),
      sum(oddsratio < 1)
      ) / sum(oddsratio > -Inf),
    p_eczema = pd_to_p(pd, verbose = FALSE)
  )

pe <- pe |>
  unnest_rvars() |>
  pivot_longer(
    cols = starts_with("oddsratio"),
    names_to = "allergy",
    names_pattern = "([a-z]*01)",
    values_to = "oddsratio"
  ) |>
  mutate(
    allergy = factor(
      allergy,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) 

pe |>
  ggplot() +
  geom_hline(yintercept = 0) +
  stat_lineribbon(
    aes(
      x = ses_mean,
      y = oddsratio),
    alpha = 0.4
  ) +
  geom_hline(yintercept = 1) +
  scale_y_continuous(
    transform = scales::pseudo_log_trans(base = 10),
    breaks = c( 1/5, 1/2, 1, 2, 5, 10, 100)
  ) +
  facet_grid(allergy ~ landscape_type, scales = "free_y") +
  labs(y = "Odds-ratio\n(green/non-green)",
       x = "Mean Socio-Economic Status")
```

```{r isaac-diversity-lt-new}
diversity_ranges <- health_data |>
  group_by(landscape_type) |>
  reframe(
    across(contains("simpson"), \(x) range(x, na.rm = TRUE))
  )

preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = "log_inverse_simpson_skin:school_context",
  conditions = make_conditions(mv_adj_mi_isaac_ses_skin, "landscape_type"),
  resp = c("eczema01", "rhinitis01", "wheezing01")
  )

preds_df_skindiv_sc_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_skindiv_sc_lt <- preds_df_skindiv_sc_lt |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_isaac_skindiv_sc_lt <- preds_df_skindiv_sc_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = exp(log_inverse_simpson_skin),
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = exp(log_inverse_simpson_skin),
      y = estimate__,
      colour = school_context
    )) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_grid(response ~ landscape_type) +
  labs(
    x = "Skin microbial diversity",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "Percentage of children sensitized (+/- 95% CI)"
  ) +
  scale_x_log10()

p_isaac_skindiv_sc_lt

saveRDS(
  p_isaac_skindiv_sc_lt,
  file.path(data_objects_folder, "p_isaac_skindiv_sc_lt.rds")
)
```

```{r calc-isaac-diversity-slopes-new}
# Obtain slopes for each landscape, school 
drw <- as_draws_rvars(mv_adj_mi_isaac_ses_skin)
slope_drw <- expand_grid(
  response = c("eczema", "rhinitis", "wheezing"),
  landscape_type = c("urban", "rural"),
  school_context = c("non-green", "green"),
  effect = c("skin")
) |> 
  mutate(
    slope = case_when(
      landscape_type == "rural"  & school_context == "non-green" &
        response == "eczema" & effect == "skin" ~
        drw$bsp_eczema01_milog_inverse_simpson_skin 
      + drw$`bsp_eczema01_milog_inverse_simpson_skin:landscape_typerural`,
      landscape_type == "rural"  & school_context == "non-green" &
        response == "rhinitis" & effect == "skin"  ~
        drw$bsp_rhinitis01_milog_inverse_simpson_skin 
      + drw$`bsp_rhinitis01_milog_inverse_simpson_skin:landscape_typerural`,
      landscape_type == "rural"  & school_context == "non-green" &
        response == "wheezing" & effect == "skin"  ~
        drw$bsp_wheezing01_milog_inverse_simpson_skin 
      + drw$`bsp_wheezing01_milog_inverse_simpson_skin:landscape_typerural`,
      landscape_type == "urban"  & school_context == "non-green" 
      & response == "eczema" & effect == "skin"  ~
        drw$bsp_eczema01_milog_inverse_simpson_skin,
      landscape_type == "urban"  & school_context == "non-green" 
      & response == "rhinitis" & effect == "skin"  ~
        drw$bsp_rhinitis01_milog_inverse_simpson_skin,
      landscape_type == "urban" & school_context == "non-green" 
      & response == "wheezing" & effect == "skin"  ~
        drw$bsp_wheezing01_milog_inverse_simpson_skin,
      landscape_type == "rural"  & school_context == "green" &
        response == "eczema" & effect == "skin" ~
        drw$bsp_eczema01_milog_inverse_simpson_skin 
      + drw$`bsp_eczema01_milog_inverse_simpson_skin:landscape_typerural`
      + drw$`bsp_eczema01_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "rural"  & school_context == "green" &
        response == "rhinitis" & effect == "skin"  ~
        drw$bsp_rhinitis01_milog_inverse_simpson_skin 
      + drw$`bsp_rhinitis01_milog_inverse_simpson_skin:landscape_typerural`
      + drw$`bsp_rhinitis01_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "rural"  & school_context == "green" &
        response == "wheezing" & effect == "skin"  ~
        drw$bsp_wheezing01_milog_inverse_simpson_skin 
      + drw$`bsp_wheezing01_milog_inverse_simpson_skin:landscape_typerural`
      + drw$`bsp_wheezing01_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "urban"  & school_context == "green" 
      & response == "eczema" & effect == "skin"  ~
        drw$bsp_eczema01_milog_inverse_simpson_skin
      + drw$`bsp_eczema01_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "urban"  & school_context == "green" 
      & response == "rhinitis" & effect == "skin"  ~
        drw$bsp_rhinitis01_milog_inverse_simpson_skin
      + drw$`bsp_rhinitis01_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "urban" & school_context == "green" 
      & response == "wheezing" & effect == "skin"  ~
        drw$bsp_wheezing01_milog_inverse_simpson_skin
      + drw$`bsp_wheezing01_milog_inverse_simpson_skin:school_contextgreen`
    )      
  )

slopes_drw <- slope_drw |> 
  rowwise() |> 
  mutate(
    median = median(slope),
    lower2.5 = quantile(slope, probs = 0.025),
    upper97.5 = quantile(slope, probs = 0.975),
    pd = pmax(
      sum(slope > 0),
      sum(slope < 0)
      ) / sum(slope > -Inf),
    p_value = pd_to_p(pd, verbose = FALSE)
  )

```

(ref:isaac-diversity-slopes-skin) Slopes (median and 95% credible intervals) for the effect of microbial skin diversity on ISAAC outcomes. Slope values are in log-odds scale and diversity values are log-transformed. Therefore, interpretation a k-fold increase of diversity is associated with a k\^slope multiplication of the odds.

```{r isaac-diversity-slopes-skin}
slopes_drw |> 
  select(-slope) |> 
  mutate(
    slope = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    ),
    ten_fold_effect = sprintf(
      "%.2f [%.2f - %.2f]",
      10 ^ median,
      10 ^ lower2.5,
      10 ^ upper97.5
    )
  ) |> 
  select(
    response, landscape_type,
    schoolyard = school_context,
    effect, slope, ten_fold_effect, p_value
  ) |> 
  mutate(
    response = factor(response, levels = c("rhinitis", "wheezing", "eczema"))
  ) |> 
  arrange(response) |> 
  kable(
    digits = 2,
    caption = "(ref:isaac-diversity-slopes-skin)"
  )

```

```{r isaac-diversity-slope-differences-skin}
slope_differences_drw <- slope_drw |>
  group_by(response, landscape_type, effect) |>
  summarise(
    slope_difference = slope[school_context == "green"] -
      slope[school_context == "non-green"],
    .groups = "drop_last"
  ) |> 
  mutate(
    difference = "green - non-green"
  ) |> 
  rowwise() |> 
  mutate(
    median = median(slope_difference),
    lower2.5 = quantile(slope_difference, probs = 0.025),
    upper97.5 = quantile(slope_difference, probs = 0.975),
    pd = pmax(
      sum(slope_difference > 0),
      sum(slope_difference < 0)
      ) / sum(slope_difference > -Inf),
    p_value = pd_to_p(pd, verbose = FALSE)
  )

slope_differences_drw |>
  select(-slope_difference) |> 
  mutate(
    slope_difference = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    )
  ) |> 
  select(response, effect, difference, slope_difference, p_value) |> 
  mutate(
    response = factor(response, levels = c("rhinitis", "wheezing", "eczema"))
  ) |> 
  arrange(response) |> 
  kable(digits = 2)
```

```{r conditional-effects-adjusters-isaac-new}
adj_1 <- c("biological_sex", "f_allergies_family")
adj_2 <- c(
  "allergenic_potential"
)

adjusters <- c(
  adj_1, adj_2
)


preds <- conditional_effects(
  mv_adj_mi_isaac_ses_skin,
  effects = adjusters,
  resp = c("eczema01", "rhinitis01", "wheezing01")
)

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    adjuster <- names(x)[1]
    x$response <- response
    x$effect1__ <- adjuster
    x
  }
)

preds_df <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()
```

```{r isaac-adjusters-new}
p_isaac_allergenic_potential <- preds_df |>
  select(-biological_sex, -f_allergies_family, -landscape_type, -school_context) |>
  pivot_longer(
    cols = any_of(adjusters),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(
    variable == effect1__,
    effect1__ %in% adj_2
  ) |>
  mutate(
    effect1__ = stringr::str_wrap(gsub("_", " ", effect1__), 15),
    effect1__ = reorder(effect1__, estimate__, FUN = \(x) diff(range(x)))
  ) |> 
  ggplot() +
  geom_ribbon(
    aes(x = value,
        y = estimate__,
        ymin = lower__,
        ymax = upper__),
    alpha = 0.4
    ) +
  geom_line(
    aes(
      x = value,
      y = estimate__
    )
  ) +
  facet_grid(response ~ effect1__,
             scales = "free") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(y = "Percentage of sensitized children (+/- 95% CI)")

p_isaac_allergenic_potential

saveRDS(
  p_isaac_allergenic_potential,
  file.path(data_objects_folder, "p_isaac_allergenic_potential.rds")
)
```

```{r isaac-biological-sex-new}
p_isaac_biological_sex <- preds_df |>
  filter(effect1__ == "biological_sex") |>
  ggplot() +
  geom_pointrange(
    aes(
      x = response,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = biological_sex 
    ),
    position = position_dodge(width = 0.2)
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of children\n(+/- 95% CI)")

p_isaac_biological_sex

saveRDS(
  p_isaac_biological_sex,
  file.path(data_objects_folder, "p_isaac_biological_sex.rds")
)
```

```{r isaac-biological-sex-riskratio}
nd <- expand.grid(
  biological_sex = c("Male", "Female")) |>
  mutate(
    school_context = "non-green",
    landscape_type = "urban",
    f_allergies_family = "no",
    ses_centered = 0,
    ses_mean = mean(health_data$ses_mean, na.rm = TRUE) |> round(2),
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |>
      round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE)|> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE)|> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    log_inverse_simpson_skin = mean(health_data$log_inverse_simpson_skin, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    allergenic_potential = mean(
      health_data$allergenic_potential, na.rm = TRUE
    ) |>
      round(2)
  )

# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_isaac_ses_skin,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("rhinitis01", "wheezing01", "eczema01")
)

# calculate risk ratio
pe <- pe |>
  select(
    biological_sex, .linpred
  ) |> 
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "riskratio"
    ),
    contrast = paste(last(biological_sex), " / ", first(biological_sex))
  )

pe |> 
   mutate(
    rr_eczema = c(riskratio[, "eczema01"]),
    rr_rhinitis = c(riskratio[, "rhinitis01"]),
    rr_wheezing = c(riskratio[, "wheezing01"])
  ) |>
  select(-riskratio) |> 
  pivot_longer(
    cols = starts_with("rr_"),
    names_to = "response",
    names_prefix = "rr_",
    values_to = "riskratio"
  ) |>
  mutate(
    pd = pmax(sum(riskratio > 1), sum(riskratio < 1)) / sum(riskratio > -Inf),
    p_value_one_sided = pd_to_p(pd, direction = "one-sided"),
    p_value_two_sided = pd_to_p(pd, direction = "two-sided"),
    riskratio = sprintf(
      "%.2f [%.2f - %.2f]",
      median(riskratio),
      quantile2(riskratio, 0.025),
      quantile2(riskratio, 0.975)
    )
  ) |> 
  select(-pd) |> 
  mutate(
    response = factor(response, levels = c("rhinitis", "wheezing", "eczema"))
  ) |>
  arrange(response) |> 
  kable(digits = 2)

```

```{r isaac-f-allergies-family-new}
p_isaac_allergies_family <- preds_df |>
  filter(effect1__ == "f_allergies_family") |>
  ggplot() +
  geom_pointrange(
    aes(
      x = response,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = f_allergies_family 
    ),
    position = position_dodge(width = 0.2)
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage of children (+/- 95% CI)")

p_isaac_allergies_family

saveRDS(
  p_isaac_allergies_family,
  file.path(data_objects_folder, "p_isaac_allergies_family.rds")
)
```

## Allergy category count (ISAAC survey data)

```{r adj-mi-nisaac, warning = FALSE}
priors <- set_prior(
  prior = "normal(0,1)",
  class = "b", 
  resp = c(
    "nsensitive",
    "loginversesimpsonskin",
    "sesmean",
    "allergenicpotential"
  )
)

bf_adj_mi_nisaac <-
  bf(
    update(rhs_isaac, n_sensitive | trials(3) ~ . + mi(allergenic_potential)),
    decomp = "QR",
    family = binomial()
  ) +
  bf(
    ses_mean | mi() + trunc(lb = 0, ub = 1) 
    ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    log_inverse_simpson_skin | mi() ~ school_context*landscape_type
    + (1|school_id),
    family = gaussian()
  ) +
  bf(
    allergenic_potential | mi() + trunc(lb = 0, ub = 4) ~
      school_context*landscape_type,
    family = gaussian()
  ) +
  set_rescor(FALSE)




mv_adj_mi_nisaac <- brm(
    formula = bf_adj_mi_nisaac,
    data = health_data,
    prior = priors,
    cores = 4,
    silent = 2,
    refresh = 0,
    chains = 4,
    backend = "cmdstanr",
    normalize = FALSE,
    file_refit = "on_change",
    file = file.path(
      model_folder,
      "mv_adj_mi_nisaac")
)

```

```{r summary-nisaac, warning=FALSE}
pvals <- pd(mv_adj_mi_nisaac) %>%
  as_tibble() %>%
  separate_wider_delim(
    Parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) %>%
  select(-dropme) %>%
  mutate(p_value = pd_to_p(pd)) %>%
  select(-Effects, -Component)

gather_draws(mv_adj_mi_nisaac, `b_nsen.*|bsp_nsen.*`,
             regex = TRUE) %>%
  separate_wider_delim(
    .variable,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) %>%
  select(-dropme) %>%
  group_by(response, parameter) %>%
  summarise_draws() %>%
  inner_join(pvals) %>%
  relocate(p_value, .before = rhat) %>%
  select(-variable, -rhat, -ess_bulk, -ess_tail) %>%
  kable(digits = 2)
```

\@ref(fig:mcmc-intervals-nisaac)

(ref:mcmc-intervals-nisaac) Parameter estimates for the effects on the Allergy category count. Point estimates represent the median of the posterior distribution. Thick and thin lines show 50% and 90% inner probability mass of the posterior distribution. Effects are with reference to urban landscape type, boys and non-green schoolyards.

```{r mcmc-intervals-nisaac, fig.height = 200/25.4, out.width="100%", fig.cap="(ref:mcmc-intervals-nisaac)"}
nisaac_alldata <- mcmc_intervals_data(
  mv_adj_mi_nisaac,
  regex_pars = "^(b|bsp)_nsensitive",
  prob = 0.5, # inner interval probability mass
  prob_outer = 0.9, # outer interval probability mass
  point_est = "median"
) |> 
  mutate(
    dataset = "all data"
  )
nisaac_alldata <- nisaac_alldata %>%
  mutate(
    response = "nsensitive",
    parameter = gsub("^(b_nsensitive_|bsp_nsensitive_)", "", parameter),
    parameter = gsub("landscape_typerural", "Rural", parameter),
    parameter = gsub("school_contextgreen", "Green schoolyard", parameter),
    parameter = gsub("biological_sexFemale", "Biological sex - girl", parameter),
    parameter = gsub("mises_mean", "Mean school SES", parameter),
    parameter = gsub("milog_inverse_simpson_skin", "Diversity skin microbiome", parameter),
    parameter = gsub("f_allergies_familyyes", "Allergies in family - yes", parameter),
    parameter = gsub("f_allergies_familyunknown", "Allergies in family - unknown", parameter),
    parameter = gsub("miallergenic_potential", "Allergenic potential", parameter),
    parameter = reorder(parameter, m)
  )

mcmc_intervals_nisaac <- nisaac_alldata |>
  filter(parameter != "Intercept") |>
  ggplot() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_linerange(
    aes(
      x = m,
      xmin = ll,
      xmax = hh,
      y = parameter
    ),
    linewidth = 1, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_linerange(
    aes(
      x = m,
      xmin = l,
      xmax = h,
      y = parameter
    ),
    linewidth = 2, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_point(
    aes(
      x = m,
      y = parameter
    ),
    size = 3, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  labs(
    x = "Parameter estimates"
  ) +
  theme(
    legend.position = "top",
    axis.title.y = element_blank()
  )

mcmc_intervals_nisaac

saveRDS(
  mcmc_intervals_nisaac,
  file.path(data_objects_folder, "mcmc_intervals_nisaac.rds")
)
```

Posterior predictive checks:

```{r ppcheck-nisaac}
responses <- mv_adj_mi_nisaac$formula$responses
checks <- vector("list", length(responses))
for (i in seq_along(responses)) {
  checks[[i]] <- pp_check(
    mv_adj_mi_nisaac,
    resp = responses[i],
    ndraws = 50) + labs(x = responses[i])
}
checks %>%
  wrap_plots()
```

```{r}
preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "landscape_type:school_context",
  resp = c("nsensitive")
  )

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_nisaac_ltsc <- preds_df |>
  as_tibble()
```

```{r adj-mi-nisaac-preds-sclt}
p_nisaac_ltsc <- preds_df_nisaac_ltsc |>
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context
    ),
    position = position_dodge(width = 0.5)
  ) +
  scale_colour_manual(values = school_colours) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Allergy category count (+/- 95% CI)"
  ) +
  scale_y_continuous(limits = c(0, 3))

p_nisaac_ltsc

saveRDS(
  p_nisaac_ltsc,
  file.path(data_objects_folder, "p_nisaac_ltsc.rds")
)

```

Figure \@ref(fig:adj-mi-nisaac-preds-sclt-school-effects)

(ref:adj-mi-nisaac-preds-sclt-school-effects) Predicted number of allergenic sensitizations for each school as a function of landscape type and schoolyard. The centrality measure is the median and the line intervals correspond to 95% credible intervals.

```{r adj-mi-nisaac-preds-sclt-school-effects, fig.cap = "(ref:adj-mi-nisaac-preds-sclt-school-effects)"}
preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "school_id",
  conditions = make_conditions(
    mv_adj_mi_nisaac, vars = c("landscape_type", "school_context")
  ),
  resp = c("nsensitive"),
  re_formula = NULL
  )

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

# filter out the actual types to which a school belongs
school_cats <- health_data |>
  distinct(school_id, landscape_type, school_context)
preds_df <- preds_df |>
  semi_join(
    school_cats,
    by = join_by(school_id, landscape_type, school_context)
  )


preds_df_nisaac_ltsc_schooleffects <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = c("nsensitive"),
      labels = c("Number of allergenic sensitizations")
    )
  ) |>
  as_tibble()

p_nisaac_ltsc_schooleffects <- preds_df_nisaac_ltsc_schooleffects |>
  # Group by the facet and x-axis variable
  group_by(response, landscape_type) |> 
  # Sort ascending by estimate__ (per your request)
  arrange(desc(estimate__)) |> 
  # Create a local rank for the dodging order
  mutate(dodge_rank = factor(row_number())) |> 
  ungroup() |> 
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context,
      # Map the group aesthetic to the local rank
      group = dodge_rank 
    ),
    size = 0.25,
    alpha = 0.5,
    position = position_dodge(width = 0.8)
  ) +
  scale_colour_manual(values = school_colours) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Allergy category counts (+/- 95% CI)"
  ) +
  scale_y_continuous(limits = c(0, 3))

p_nisaac_ltsc_schooleffects

saveRDS(
  p_nisaac_ltsc_schooleffects,
  file.path(data_objects_folder, "p_nisaac_ltsc_schooleffects.rds")
)
```

We can calculate post-hoc comparison, stratified by landscape type to see if school types differ significantly within landscape types:

```{r emmeans-sclt-nsens}
#Error : brms' emmeans support for multivariate models is limited to cases where all univariate models have the same family.
# instead calc manually
nd <- expand.grid(
  school_context = c("non-green", "green"),
  ses_mean = median(mv_adj_mi_nisaac$data$ses_mean, na.rm = TRUE),
  landscape_type = c("urban", "rural")
) %>%
  mutate(
    biological_sex = "Male",
    ses_centered = 0,
    f_allergies_family = "no",
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |>
      round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE)|> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE)|> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    log_inverse_simpson_skin = mean(health_data$log_inverse_simpson_skin, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    allergenic_potential = mean(
      health_data$allergenic_potential, na.rm = TRUE
    ) |>
      round(2)
  ) %>%
  inner_join(ses_ranges_landscape) %>%
  filter(
    ses_mean >= min_overlap_ses_mean,
    ses_mean <= max_overlap_ses_mean
  )

# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_nisaac,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("nsensitive")
)

# calculate risk ratio
pe <- pe %>%
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "riskratio"
    ),
    contrast = paste(last(school_context), "/", first(school_context)),
    .by = c(ses_mean, landscape_type, biological_sex, ses_centered,
            garden_score, natural_elements_score, nature_connectedness,
            mean_log_sirm, log_invsimpson_plant,
            log_invsimpson_dust,
            log_invsimpson_soil,
            log_inverse_simpson_skin,
            sum_contact_animals,
            f_allergies_family,
            allergenic_potential)
  ) 
pe_nisaac_contrasts_sc <- pe |> 
  select(landscape_type, contrast, riskratio) |>
  mutate(
    pd = pmax(sum(riskratio > 1), sum(riskratio < 1)) / sum(riskratio > -Inf),
    p_value_two_sided = pd_to_p(pd),
    p_value_one_sided = pd_to_p(pd, direction = "one-sided"),
    riskratio = sprintf(
      "%.2f [%.2f - %.2f]",
      median(riskratio),
      quantile2(riskratio, 0.025),
      quantile2(riskratio, 0.975)
    )
  ) |> 
  select(-pd) 
pe_nisaac_contrasts_sc |> 
  kable(digits = 2)

```

Figure \@ref(fig:sei-sc-interaction-nisaac) shows the interaction between school-level SES, on the one hand, and Schoolyard or landscape type, on the other hand.

(ref:sei-sc-interaction-nisaac) Two-way interaction conditional effects between Schoolyard and landscape type, school-level socio-economic index and Schoolyard, school-level socio-economic index and landscape type, on number sensitized (backtransformed to original scale). The median of the posterior distribution and 95% credible intervals are shown.

```{r sei-sc-interaction-nisaac, fig.height=120/25.4, out.width="110%", fig.cap="(ref:sei-sc-interaction-nisaac)"}
preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "ses_mean:school_context",
  resp = "nsensitive"
  )

preds_df_ses_sc <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_sc <- preds_df_ses_sc %>%
  inner_join(ses_ranges) %>%
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_sc <- preds_df_ses_sc |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

p_nisaac_ses_sc <- preds_df_ses_sc |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = school_context
    )) +
  scale_y_continuous(limits = c(0, 3)) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "Allergy category count (+/- 95% CI)"
  )

preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "ses_mean:landscape_type",
  resp = "nsensitive"
  )

preds_df_ses_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_lt <- preds_df_ses_lt %>%
  inner_join(ses_ranges) %>%
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  )
  
preds_df_ses_lt <- preds_df_ses_lt |>
  mutate(
    response = factor(
      response,
      levels = kidscreen_groups$level,
      labels = kidscreen_groups$short_label
    )
  ) |>
  as_tibble()

p_nisaac_ses_lt <- preds_df_ses_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = landscape_type
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = landscape_type
    )) +
  scale_y_continuous(limits = c(0, 3)) +
  scale_colour_manual(
    values = landscape_colours) +
  scale_fill_manual(
    values = landscape_colours) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Landscape type",
    fill = "Landscape type",
    y = "Allergy category count (+/- 95% CI)"
  )

p_nisaac_ltsc + p_nisaac_ses_sc + p_nisaac_ses_lt +
  plot_layout(guides = "collect", axes = "collect") +
  plot_annotation(tag_levels = "a")

saveRDS(
  p_nisaac_ses_sc,
  file.path(data_objects_folder, "p_nisaac_ses_sc.rds")
)
saveRDS(
  p_nisaac_ses_lt,
  file.path(data_objects_folder, "p_nisaac_ses_lt.rds")
)

```

Alternative visualisation that better shows how the three two-way interactions combine:

```{r}
preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "ses_mean:school_context",
  conditions = make_conditions(mv_adj_mi_nisaac, "landscape_type"),
  resp = "nsensitive"
  )

preds_df_ses_sc_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_ses_sc_lt <- preds_df_ses_sc_lt %>%
  inner_join(ses_ranges) %>%
  filter(
    ses_mean >= min_ses_mean,
    ses_mean <= max_ses_mean
  ) |>
  as_tibble()

p_nisaac_ses_sc_lt <- preds_df_ses_sc_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = ses_mean,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = ses_mean,
      y = estimate__,
      colour = school_context
    )) +
  scale_y_continuous(limits = c(0, 3)) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_wrap(~ landscape_type) +
  labs(
    x = "School-level Socio-Economic Index",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "Allergy category count (+/- 95% CI)"
  )

p_nisaac_ses_sc_lt

saveRDS(
  p_nisaac_ses_sc_lt,
  file.path(data_objects_folder, "p_nisaac_ses_sc_lt.rds")
)
```

The corresponding slopes are:

```{r nisaac-slopes-2way}
# Obtain slopes for each landscape, school 
drw <- as_draws_rvars(mv_adj_mi_nisaac)
slope_drw <- list(
  green_rural = drw$bsp_nsensitive_mises_mean 
  + drw$`bsp_nsensitive_mises_mean:landscape_typerural`
  + drw$`bsp_nsensitive_mises_mean:school_contextgreen`,
  green_urban = drw$bsp_nsensitive_mises_mean 
  + drw$`bsp_nsensitive_mises_mean:school_contextgreen`,
  nongreen_rural = drw$bsp_nsensitive_mises_mean 
  + drw$`bsp_nsensitive_mises_mean:landscape_typerural`,
  nongreen_urban = drw$bsp_nsensitive_mises_mean
)

parameter <- names(slope_drw)
rvars <- map_vec(slope_drw, \(x) x[[1]])
mcmc_df <- tibble(
  parameter = parameter,
  slope = rvars
) |> 
  rowwise() |> 
  mutate(
    median = median(slope),
    lower2.5 = quantile(slope, probs = 0.025),
    upper97.5 = quantile(slope, probs = 0.975),
    pd = pmax(
      sum(slope > 0),
      sum(slope < 0)
      ) / sum(slope > -Inf),
    p_value = pd_to_p(pd)
  )

mcmc_df |> 
  mutate(
    slope = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    )) |> 
  select(parameter, slope, p_value) |> 
kable(digits = 2)
```

```{r nisaac-diversity-lt-new}
diversity_ranges <- health_data |>
  group_by(landscape_type) |>
  reframe(
    across(contains("simpson"), \(x) range(x, na.rm = TRUE))
  )

preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = "log_inverse_simpson_skin:school_context",
  conditions = make_conditions(mv_adj_mi_nisaac, "landscape_type"),
  resp = c("nsensitive")
  )

preds_df_skindiv_sc_lt <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df_skindiv_sc_lt <- preds_df_skindiv_sc_lt %>%
  as_tibble()

p_nisaac_skindiv_sc_lt <- preds_df_skindiv_sc_lt |>
  ggplot() +
  geom_ribbon(
    aes(
      x = exp(log_inverse_simpson_skin),
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      fill = school_context
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = exp(log_inverse_simpson_skin),
      y = estimate__,
      colour = school_context
    )) +
  scale_y_continuous(limits = c(0, 3)) +
  scale_colour_manual(
    values = school_colours) +
  scale_fill_manual(
    values = school_colours) +
  facet_grid(~ landscape_type) +
  labs(
    x = "Skin microbial diversity",
    colour = "Schoolyard",
    fill = "Schoolyard",
    y = "Predicted Allergy category count (+/- 95% CI)"
  ) +
  scale_x_log10()

p_nisaac_skindiv_sc_lt

saveRDS(
  p_nisaac_skindiv_sc_lt,
  file.path(data_objects_folder, "p_nisaac_skindiv_sc_lt.rds")
)
```

```{r calc-nisaac-diversity-slopes-new}
# Obtain slopes for each landscape, school 
drw <- as_draws_rvars(mv_adj_mi_nisaac)
slope_drw <- expand_grid(
  response = c("nsensitive"),
  landscape_type = c("urban", "rural"),
  school_context = c("non-green", "green"),
  effect = c("skin")
) |> 
  mutate(
    slope = case_when(
      landscape_type == "rural"  & school_context == "non-green"
      & effect == "skin" ~
        drw$bsp_nsensitive_milog_inverse_simpson_skin 
      + drw$`bsp_nsensitive_milog_inverse_simpson_skin:landscape_typerural`,
      landscape_type == "urban"  & school_context == "non-green" &
        effect == "skin"  ~
        drw$bsp_nsensitive_milog_inverse_simpson_skin,
      landscape_type == "rural"  & school_context == "green" &
        effect == "skin" ~
        drw$bsp_nsensitive_milog_inverse_simpson_skin 
      + drw$`bsp_nsensitive_milog_inverse_simpson_skin:landscape_typerural`
      + drw$`bsp_nsensitive_milog_inverse_simpson_skin:school_contextgreen`,
      landscape_type == "urban"  & school_context == "green" 
      & effect == "skin"  ~
        drw$bsp_nsensitive_milog_inverse_simpson_skin
      + drw$`bsp_nsensitive_milog_inverse_simpson_skin:school_contextgreen`
    )      
  )

slopes_drw <- slope_drw |> 
  rowwise() |> 
  mutate(
    median = median(slope),
    lower2.5 = quantile(slope, probs = 0.025),
    upper97.5 = quantile(slope, probs = 0.975),
    pd = pmax(
      sum(slope > 0),
      sum(slope < 0)
      ) / sum(slope > -Inf),
    p_value = pd_to_p(pd)
  )

```

(ref:nisaac-diversity-slopes-skin) Slopes (median and 95% credible intervals) for the effect of microbial skin diversity on ISAAC outcomes. Slope values are in log-odds scale and diversity values are log-transformed. Therefore, interpretation a k-fold increase of diversity is associated with a k\^slope multiplication of the odds. For instance, a 10 fold increase in skin microbial diversity in an urban, green schoolyard setting has the effect of reducing the odds of the Allergy category count by a factor $10^{-0.40} \approx \times 2/5$.

```{r nisaac-diversity-slopes-skin}
slopes_drw |> 
  select(-slope) |> 
  mutate(
    slope = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    ),
    ten_fold_effect = sprintf(
      "%.2f [%.2f - %.2f]",
      10 ^ median,
      10 ^ lower2.5,
      10 ^ upper97.5
    )
  ) |> 
  select(
    response, landscape_type,
    schoolyard = school_context,
    effect, slope, ten_fold_effect, p_value
  ) |> 
  arrange(response) |> 
  kable(
    digits = 2,
    caption = "(ref:nisaac-diversity-slopes-skin)"
  )

```

```{r nisaac-diversity-slope-differences-skin}
slope_differences_drw <- slope_drw |>
  group_by(response, landscape_type, effect) |>
  summarise(
    slope_difference = slope[school_context == "green"] - slope[school_context == "non-green"]
  ) |> 
  mutate(
    difference = "green - non-green"
  ) |> 
  rowwise() |> 
  mutate(
    median = median(slope_difference),
    lower2.5 = quantile(slope_difference, probs = 0.025),
    upper97.5 = quantile(slope_difference, probs = 0.975),
    pd = pmax(
      sum(slope_difference > 0),
      sum(slope_difference < 0)
      ) / sum(slope_difference > -Inf),
    p_value = pd_to_p(pd)
  )

slope_differences_drw |>
  select(-slope_difference) |> 
  mutate(
    slope_difference = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      lower2.5,
      upper97.5
    )
  ) |> 
  select(response, effect, difference, slope_difference, p_value) |> 
  kable(digits = 2)
```

```{r conditional-effects-adjusters-nisaac}
preds <- conditional_effects(
  mv_adj_mi_nisaac,
  effects = adjusters,
  resp = c("nsensitive")
)

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    adjuster <- names(x)[1]
    x$response <- response
    x$effect1__ <- adjuster
    x
  }
)
```

```{r}
p_nisaac_allergenic_potential <- preds_df %>%
  select(-biological_sex, -f_allergies_family, -landscape_type, -school_context) %>%
  pivot_longer(
    cols = any_of(adjusters),
    names_to = "variable",
    values_to = "value"
  ) %>%
  filter(
    variable == effect1__,
    effect1__ %in% adj_2
  ) %>%
  ggplot() +
  geom_ribbon(
    aes(x = value,
        y = estimate__,
        ymin = lower__,
        ymax = upper__),
    alpha = 0.4
    ) +
  geom_line(
    aes(
      x = value,
      y = estimate__
    )
  ) +
  scale_y_continuous(limits = c(0, 3)) +
  labs(y = "Predicted Allergy category count (+/- 95% CI)",
       x = "Allergenic potential of schoolyard")

p_nisaac_allergenic_potential
saveRDS(
  p_nisaac_allergenic_potential,
  file.path(data_objects_folder, "p_nisaac_allergenic_potential.rds")
)
```

```{r nisaac-biological-sex, fig.width = 120/25.4}
p_nisaac_biological_sex <- preds_df %>%
  filter(effect1__ == "biological_sex") %>%
  ggplot() +
  geom_pointrange(
    aes(
      x = biological_sex,
      y = estimate__,
      ymin = lower__,
      ymax = upper__
    ),
    position = position_dodge(width = 0.2)
  ) +
  scale_y_continuous(limits = c(0, 3)) +
  labs(y = "Predicted Allergy category count (+/- 95% CI)",
       x = "") +
  theme(axis.title.x = element_blank())

p_nisaac_biological_sex

saveRDS(
  p_nisaac_biological_sex,
  file.path(data_objects_folder, "p_nisaac_biological_sex.rds")
)
```

```{r nisaac-biological-sex-risk-ratio}
nd <- expand.grid(
  biological_sex = c("Male", "Female")) %>%
  mutate(
    school_context = "non-green",
    landscape_type = "urban",
    f_allergies_family = "no",
    ses_centered = 0,
    ses_mean = mean(health_data$ses_mean, na.rm = TRUE) |> round(2),
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |>
      round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE)|> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE)|> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    log_inverse_simpson_skin = mean(health_data$log_inverse_simpson_skin, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    allergenic_potential = mean(
      health_data$allergenic_potential, na.rm = TRUE
    ) |>
      round(2)
  )

# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_nisaac,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("nsensitive")
)

# calculate risk ratio
pe <- pe |>
  select(
    biological_sex, .linpred
  ) |> 
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "riskratio"
    ),
    contrast = paste(last(biological_sex), " / ", first(biological_sex))
  )

pe %>%
  mutate(
    pd = pmax(sum(riskratio > 1), sum(riskratio < 1)) / sum(riskratio > -Inf),
    p_value_one_sided = pd_to_p(pd, direction = "one-sided"),
    p_value_two_sided = pd_to_p(pd, direction = "two-sided"),
    riskratio = sprintf(
      "%.2f [%.2f - %.2f]",
      median(riskratio),
      quantile2(riskratio, 0.025),
      quantile2(riskratio, 0.975)
    )
  ) |> 
  select(-pd) |> 
  kable(digits = 2)
```

```{r nisaac-f-allergies-family-new}
p_nisaac_allergies_family <- preds_df %>%
  filter(effect1__ == "f_allergies_family") %>%
  ggplot() +
  geom_pointrange(
    aes(
      x = f_allergies_family,
      y = estimate__,
      ymin = lower__,
      ymax = upper__
    ),
    position = position_dodge(width = 0.2)
  ) +
  scale_y_continuous(limits = c(0, 3)) +
  labs(y = "Predicted Allergy category count (+/- 95% CI)")

p_nisaac_allergies_family

saveRDS(
  p_nisaac_allergies_family,
  file.path(data_objects_folder, "p_nisaac_allergies_family.rds")
)
```

## Combined figures

```{r fig-main-1, fig.height=140/25.4, fig.width = 180/25.4}
pe_isaac_contrasts_sc <- pe_isaac_contrasts_sc |> 
  mutate(
    star = case_when(
      p_value_one_sided <= 0.05 ~ "*",
      p_value_one_sided <= 0.1 ~ "^",
      TRUE ~ ""
    ),
    label = sprintf("p = %.02f%s", p_value_one_sided, star)
  )
pe_nisaac_contrasts_sc <- pe_nisaac_contrasts_sc |> 
  mutate(
    star = case_when(
      p_value_one_sided <= 0.05 ~ "*",
      p_value_one_sided <= 0.1 ~ "^",
      TRUE ~ ""
    ),
    label = sprintf("p = %.02f%s", p_value_one_sided, star)
  )

p_isaac_ltsc_combined <- preds_df_isaac_ltsc_schooleffects |>
  # Group by the facet and x-axis variable
  group_by(response, landscape_type) |> 
  mutate(
    sign = mean(estimate__[school_context == "green"]) -
      mean(estimate__[school_context == "non-green"]),
    sortorder = sign * estimate__
  ) |> 
  # Sort by sign * estimate__
  arrange(sortorder) |> 
  # Create a local rank for the dodging order
  mutate(dodge_rank = factor(row_number())) |> 
  ungroup() |> 
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context,
      # Map the group aesthetic to the local rank
      group = dodge_rank 
    ),
    linewidth = 0.2,
    size = 0.2,
    alpha = 0.4,
    position = position_dodge(width = 0.8)
  ) +
  geom_pointrange(
    data = preds_df_isaac_ltsc,
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context
    ),
    position = position_dodge(width = 0.8)
  ) +
  geom_text(
    data = pe_isaac_contrasts_sc,
    aes(
      x = landscape_type,
      y = c(0.85, 0.75, 0.5, 0.4, 0.5, 0.25),
      label = label
    )
  ) + 
  scale_colour_manual(values = school_colours) +
  facet_wrap(~ response) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Percentage of sensitized children\n(+/- 95% CI)"
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA))


p_nisaac_ltsc_combined <- preds_df_nisaac_ltsc_schooleffects |>
  # Group by the facet and x-axis variable
  group_by(response, landscape_type) |> 
  mutate(
    sign = mean(estimate__[school_context == "green"]) -
      mean(estimate__[school_context == "non-green"]),
    sortorder = sign * estimate__
  ) |> 
  # Sort by sign * estimate__
  arrange(sortorder) |> 
  # Create a local rank for the dodging order
  mutate(dodge_rank = factor(row_number())) |> 
  ungroup() |> 
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context,
      # Map the group aesthetic to the local rank
      group = dodge_rank 
    ),
    linewidth = 0.2,
    size = 0.2,
    alpha = 0.4,
    position = position_dodge(width = 0.8)
  ) +
  geom_pointrange(
    data = preds_df_nisaac_ltsc,
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context
    ),
    position = position_dodge(width = 0.8)
  ) +
  geom_text(
    data = pe_nisaac_contrasts_sc,
    aes(
      x = landscape_type,
      y = c(1.5, 1.2),
      label = label
    )
  ) + 
  scale_colour_manual(values = school_colours) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Allergy categories counts\n(+/- 95% CI)"
  ) +
  scale_y_continuous(limits = c(0, 3))


p_fig1 <- mcmc_intervals_kidscreen_all_data / 
  (p_isaac_ltsc_combined + 
  p_nisaac_ltsc_combined +
  plot_layout(
    design = "
    aaab
    ",
    guides = "collect", axes = "collect")) +
  plot_annotation(tag_levels = "A")

saveRDS(
  p_fig1,
  file.path(data_objects_folder, "p_fig1.rds")
)
ggsave(
  plot = p_fig1,
  filename = file.path(
    data_objects_folder, "p_fig1.png"
  ),
  width = 30, height = 25, units = "cm", bg = "white",
  dpi = 320
)
knitr::include_graphics(
  file.path(
    data_objects_folder, "p_fig1.png"
  ),
  rel_path = FALSE
)
```

## Skin microbial diversity model

```{r skindiv-model}
skin_equation <-
  inverse_simpson_skin | mi() ~
  mi(log_total_count_skin) +
  school_context +
  landscape_type +
  school_context:landscape_type +
  mi(log_invsimpson_dust) +
  mi(log_invsimpson_plant) +
  mi(log_invsimpson_soil) +
  mi(nature_connectedness) +
  mi(garden_score) +
#  mi(garden_score):landscape_type +
  mi(sum_contact_animals) +
  natural_elements_score +
  mi(mean_log_sirm) +
  (1|q|school_id)

priors_skin <- set_prior(
  prior = "normal(0,1)", class = "b", 
  resp = c(
    "gardenscore", 
    "meanlogsirm", "natureconnectedness", "inversesimpsonskin", 
    "loginvsimpsondust", "loginvsimpsonplant", "loginvsimpsonsoil", 
    "sumcontactanimals", "logtotalcountskin")
)

bf_skin <-
  bf(
    skin_equation,
    decomp = "QR",
    family = Gamma(link = "log")
  ) +
  bf(
    nature_connectedness | mi() ~ school_context*landscape_type + (1|school_id),
    family = gaussian()
  ) +
  bf(
    garden_score | mi() ~ school_context*landscape_type + (1|school_id),
    family = gaussian()
  ) +
  bf(
    mean_log_sirm | mi() + trunc(lb = 0) ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    log_invsimpson_plant | mi() ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    log_invsimpson_dust | mi() ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    log_invsimpson_soil | mi() ~ school_context*landscape_type,
    family = gaussian()
  ) +
  bf(
    sum_contact_animals | mi() ~ school_context*landscape_type
    + (1|school_id),
    family = gaussian()
  ) +
  bf(
    log_total_count_skin | mi() ~ school_context*landscape_type
    + (1|school_id),
    family = gaussian()
  ) +
  set_rescor(FALSE)


mv_adj_mi_skin <- brm(
    formula = bf_skin,
    data = health_data,
    prior = priors_skin,
    cores = 4,
    silent = 2,
    refresh = 0,
    chains = 4,
    backend = "cmdstanr",
    normalize = FALSE,
    file_refit = "on_change",
    file = file.path(
      model_folder,
      "mv_adj_mi_skin")
)
```

```{r summary-skindiv, warning=FALSE}
pvals <- pd(mv_adj_mi_skin) |>
  as_tibble() |>
  separate_wider_delim(
    Parameter,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  mutate(p_value = pd_to_p(pd, verbose = FALSE)) |>
  select(-Effects, -Component)

gather_draws(
  mv_adj_mi_skin,
  `b_(inversesimpsonskin).*|bsp_(inversesimpsonskin).*`,
  regex = TRUE
) |>
  separate_wider_delim(
    .variable,
    delim = "_",
    too_many = "merge",
    names = c("dropme", "response", "parameter")) |>
  select(-dropme) |>
  group_by(response, parameter) |>
  summarise_draws() |>
  inner_join(pvals) |>
  relocate(p_value, .before = rhat) |>
  select(-variable, -rhat, -ess_bulk, -ess_tail) |>
  mutate(
    estimate = sprintf(
      "%.2f [%.2f - %.2f]",
      median,
      q5,
      q95
    )
  ) |> 
  select(response, parameter, estimate, p_value) |>
  kable(digits = 2)
```

\@ref(fig:mcmc-intervals-skindiv)

(ref:mcmc-intervals-skindiv) Parameter estimates for the effects on microbial skin diversity. Point estimates represent the median of the posterior distribution. Thick and thin lines show 50% and 90% inner probability mass of the posterior distribution. Effects are with reference to urban landscape type and non-green schoolyards.

```{r mcmc-intervals-skindiv, fig.height = 200/25.4, out.width="120%", fig.cap="(ref:mcmc-intervals-skindiv)"}
isaac_alldata <- mcmc_intervals_data(
  mv_adj_mi_skin,
  regex_pars = "^(b|bsp)_(inversesimpsonskin)",
  prob = 0.5, # inner interval probability mass
  prob_outer = 0.9, # outer interval probability mass
  point_est = "median"
) |> 
  mutate(
    dataset = "all data"
  )

isaac_alldata <- isaac_alldata |> 
  mutate(
    response = stringr::str_extract(string = parameter, "^(b|bsp)_(.+)_", group = 2),
    parameter = gsub("^(b|bsp)_inversesimpsonskin_", "", parameter),
    parameter = gsub("landscape_typerural", "Rural", parameter),
    parameter = gsub("school_contextgreen", "Green schoolyard", parameter),
    parameter = gsub("natural_elements_score", "Natural elements score", parameter),
    parameter = gsub("migarden_score", "Garden score", parameter),
    parameter = gsub("mimean_log_sirm", "Local air pollution", parameter),
    parameter = gsub("minature_connectedness", "Nature connectedness", parameter),
    parameter = gsub("misum_contact_animals", "Contact animals", parameter),
    parameter = gsub("milog_invsimpson_plant", "Diversity plant microbiome", parameter),
    parameter = gsub("milog_invsimpson_dust", "Diversity dust microbiome", parameter),
    parameter = gsub("milog_invsimpson_soil", "Diversity soil microbiome", parameter),
    parameter = gsub("milog_total_count_skin", "Skin total read count", parameter),
    parameter = reorder(parameter, m)
  )

mcmc_skindiv <- isaac_alldata |>
  filter(
    parameter != "Intercept"
  ) |> 
  ggplot() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_linerange(
    aes(
      x = m,
      xmin = ll,
      xmax = hh,
      y = parameter
    ),
    linewidth = 1, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_linerange(
    aes(
      x = m,
      xmin = l,
      xmax = h,
      y = parameter
    ),
    linewidth = 2, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  geom_point(
    aes(
      x = m,
      y = parameter
    ),
    size = 3, alpha = 0.5, position = position_dodge(width = 0.5)
  ) +
  labs(
    x = "Parameter estimates"
  ) +
  theme(
    legend.position = "top",
    axis.title.y = element_blank()
  )

mcmc_skindiv

saveRDS(
  mcmc_skindiv,
  file.path(data_objects_folder, "mcmc_skindiv.rds")
)

```

Posterior predictive checks:

```{r ppcheck-skin-new, warning=FALSE}
responses <- mv_adj_mi_skin$formula$responses
checks <- vector("list", length(responses))
for (i in seq_along(responses)) {
  checks[[i]] <- pp_check(
    mv_adj_mi_skin,
    resp = responses[i],
    ndraws = 50) + labs(x = responses[i])
}
checks |>
  wrap_plots()
```

```{r adj-mi-skin-sclt}
preds <- conditional_effects(
  mv_adj_mi_skin,
  effects = "landscape_type:school_context",
  resp = c("inversesimpsonskin")
  )

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df <- preds_df |>
  mutate(
    response = factor(
      response,
      levels = c("rhinitis01", "wheezing01", "eczema01"),
      labels = c("rhinitis", "wheezing", "eczema")
    )
  ) |>
  as_tibble()

p_skindiv_ltsc <- preds_df |>
  ggplot() +
  geom_pointrange(
    aes(
      x = landscape_type,
      y = estimate__,
      ymin = lower__,
      ymax = upper__,
      colour = school_context
    ),
    position = position_dodge(width = 0.5)
  ) +
  scale_colour_manual(values = school_colours) +
  labs(
    x = "Landscape type",
    colour = "Schoolyard",
    y = "Microbial skin diversity (+/- 95% CI)"
  ) +
  scale_y_continuous()

p_skindiv_ltsc

saveRDS(
  p_skindiv_ltsc,
  file.path(data_objects_folder, "p_skindiv_ltsc.rds")
)
```

We can calculate post-hoc comparison, stratified by landscape type to see if school types differ significantly within landscape types:

```{r emmeans-sclt-skindiv}
#Error : brms' emmeans support for multivariate models is limited to cases where all univariate models have the same family.
# instead calc manually
nd <- expand.grid(
  school_context = c("non-green", "green"),
  landscape_type = c("urban", "rural")
) |>
  mutate(
    garden_score = mean(health_data$garden_score, na.rm = TRUE) |> round(2),
    natural_elements_score = mean(health_data$natural_elements_score) |>
      round(2),
    nature_connectedness =
      mean(health_data$nature_connectedness, na.rm = TRUE)|> round(2),
    mean_log_sirm = mean(health_data$mean_log_sirm, na.rm = TRUE)|> round(2),
    log_invsimpson_plant = mean(health_data$log_invsimpson_plant, na.rm = TRUE),
    log_invsimpson_dust = mean(health_data$log_invsimpson_dust, na.rm = TRUE),
    log_invsimpson_soil = mean(health_data$log_invsimpson_soil, na.rm = TRUE),
    sum_contact_animals = mean(health_data$sum_contact_animals, na.rm = TRUE) |>
      round(2),
    log_total_count_skin = mean(health_data$log_total_count_skin, na.rm = TRUE) |>
      round()
  )
# linear predictor link scale (logit)
pe <- tidybayes::add_linpred_rvars(
  object = mv_adj_mi_skin,
  re_formula = NA,
  newdata = nd,
  prob = 0.95,
  robust = TRUE,
  transform = FALSE,
  resp = c("inversesimpsonskin")
)

pe <- pe |>
  select(
    landscape_type, school_context, .linpred
  ) |> 
  summarise(
    across(
      starts_with(".linpred"),
      \(x) exp(last(x) - first(x)),
      .names = "ratio"
    ),
    contrast = paste(last(school_context), " / ", first(school_context)),
    .by = c(landscape_type)
  )

pe |> 
  select(landscape_type, contrast, ratio) |> 
  mutate(
    pd = pmax(sum(ratio > 1), sum(ratio < 1)) / sum(ratio > -Inf),
    p_value_one_sided = pd_to_p(pd, direction = "one-sided"),
    p_value_two_sided = pd_to_p(pd, direction = "two-sided"),
    ratio = sprintf(
      "%.2f [%.2f - %.2f]",
      median(ratio),
      quantile2(ratio, 0.025),
      quantile2(ratio, 0.975)
    )
  ) |> 
  select(-pd) |> 
  kable(digits = 2)

```

```{r skindiv-other-effects}
effects = c(
    "log_invsimpson_dust",
    "log_invsimpson_plant",
    "log_invsimpson_soil",
    "nature_connectedness",
    "garden_score",
    "sum_contact_animals",
    "natural_elements_score",
    "mean_log_sirm"
  )
preds <- conditional_effects(
  mv_adj_mi_skin,
  effects = effects,
  resp = "inversesimpsonskin"
)

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

preds_df <- preds_df |>
  as_tibble() |> 
  pivot_longer(
    cols = any_of(effects),
    names_to = "variable",
    values_to = "value"
  ) |>
  filter(
    value == effect1__
  )

p_skindiv_other_effects <- preds_df |>
  mutate(
    variable = recode_values(
      variable,
      "log_invsimpson_dust" ~ "Microbial diversity dust",
      "log_invsimpson_plant" ~ "Microbial diversity plant",
      "log_invsimpson_soil" ~ "Microbial diversity soil",
      "nature_connectedness" ~ "Nature connectedness",
      "garden_score" ~ "Garden score",
      "sum_contact_animals" ~ "Contact animals",
      "natural_elements_score" ~ "Natural elements score",
      "mean_log_sirm" ~ "SIRM"
    )
  ) |>
  ggplot() +
  geom_ribbon(
    aes(x = value,
        y = estimate__,
        ymin = lower__,
        ymax = upper__
        ),
    alpha = 0.3
    ) +
  geom_line(
    aes(
      x = value,
      y = estimate__
    )
  ) +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    y = "Microbial skin diversity"
  )

p_skindiv_other_effects

saveRDS(
  p_skindiv_other_effects,
  file.path(data_objects_folder, "p_skindiv_other_effects.rds")
)
```

```{r skin-total-count}
preds <- conditional_effects(
  mv_adj_mi_skin,
  effects = c("log_total_count_skin"),
  resp = "inversesimpsonskin"
)

preds_df <- map2_dfr(
  preds,
  names(preds),
  function(x, y) {
    response <- stringr::str_extract(y, "^\\w*")
    x$response <- response
    x
  }
)

p_skindiv_totalreads <- preds_df |> 
  ggplot() +
  geom_ribbon(
    aes(x = exp(log_total_count_skin),
        y = estimate__,
        ymin = lower__,
        ymax = upper__
    ),
    alpha = 0.3
  ) +
  geom_line(
    aes(
      x = exp(log_total_count_skin),
      y = estimate__
    )
  ) +
  scale_x_log10() +
  labs(
    x = "Skin total read count",
    y = "Skin microbial diversity"
  )


p_skindiv_totalreads

saveRDS(
  p_skindiv_totalreads,
  file.path(data_objects_folder, "p_skindiv_totalreads.rds")
)
```

# Session info

<details>

<summary>Click to view Session Info</summary>

```{r session-info, echo=FALSE}
sessioninfo::session_info()
```

</details>

# References
